---
title: "scRNA-Seq"
author: "Zuzana"
date: 
output: html_document
editor_options: 
  chunk_output_type: console
---

# Packages
Load required package

```{r, message=FALSE, warning=FALSE, paged.print=TRUE}
suppressMessages(library("scater"))
suppressMessages(library("Matrix"))
suppressMessages(library("NormExpression"))
suppressMessages(library("DropletUtils"))
suppressMessages(library("DESeq2"))
suppressMessages(library("pheatmap"))
suppressMessages(library("RColorBrewer"))
suppressMessages(library("ggplot2"))
suppressMessages(library("Seurat"))
suppressMessages(library("sctransform"))
suppressMessages(library("DT"))
suppressMessages(library("scales"))
suppressMessages(library("ggrepel"))
suppressMessages(library("dplyr"))
suppressMessages(library("SeuratWrappers"))
suppressMessages(library("scmap"))
suppressMessages(library("DoubletFinder"))
suppressMessages(library("scCATCH"))
suppressMessages(library("reshape2"))
suppressMessages(library("Nebulosa"))
library("SeuratObject")
library("grid")
library("gridExtra")
library("dittoSeq")
library("clustree")
library("pals")
library("EnhancedVolcano")
library("ggpubr")
library("clusterProfiler")
```

```{r}
samples<-read.table("samples.csv",header=TRUE,sep=",")
```

# STAR
```{r load_data, eval=FALSE, message = FALSE}
STAR<-as.list(NULL)
for(i in 1:nrow(samples)) {
  folder <- paste0("../STAR/",samples$SampleName[i],"/",samples$SampleName[i],"Solo.out/Gene/raw/")
  dims <- readr::read_delim(paste0(folder, "matrix.mtx"), delim = ' ', skip = 2, n_max = 1, col_names = FALSE)
  t <- readr::read_delim(paste0(folder, "matrix.mtx"), delim = ' ', skip = 3, col_names = FALSE)
  
  cellbarcodes <- read.table(paste0(folder, "barcodes.tsv"))
  genenames <- read.table(paste0(folder, "genes.tsv"))
  dimnames <- as.list(NULL)
  dimnames[[1]] <- as.character(genenames$V2)
  dimnames[[2]] <- as.character(paste0(cellbarcodes$V1,"-",samples$SampleName[i]))
  
  STAR[[i]]<-as.list(NULL)
  STAR[[i]][["RNA"]] <- Matrix::sparseMatrix(i = t$X1, j = t$X2, x = t$X3, dims = c(dims$X1, dims$X2), dimnames = dimnames)
}
```

```{r, eval = FALSE}
rm(cellbarcodes, dimnames, dims, genenames, t, i, folder)
saveRDS(STAR,"STAR_raw.rds")
```

# EmptyDrops

```{r, include = FALSE}
STAR <- readRDS("STAR_raw.rds")
```

```{r eval=FALSE}
p <- as.list(NULL)
p_2 <- as.list(NULL)
STAR.ED <- STAR

pdf("knee_plots.pdf")
for(i in 1:nrow(samples)) {
  bcrank = barcodeRanks(STAR[[i]]$RNA, lower = 1000)
  uniq = !duplicated(bcrank$rank)

  knee_count <- paste0("Knee = ",table(bcrank$total >= metadata(bcrank)$knee)[2]," cells")
  inflection_count <- paste0("Inflection = ",table(bcrank$total >= metadata(bcrank)$inflection)[2]," cells")
  
  p[[i]] <- qplot(bcrank$rank[uniq], bcrank$total[uniq], geom = "line") +
    geom_hline(yintercept = metadata(bcrank)$knee, color = "blue", linetype = 2) +
    geom_hline(yintercept = metadata(bcrank)$inflection, color = "darkgreen", linetype = 2) +
    annotate("text", x = 10000, y = 1.5 * c(metadata(bcrank)$knee, metadata(bcrank)$inflection),
           label = c(knee_count, inflection_count), color = c("blue", "darkgreen")) +
    scale_x_log10() +
    scale_y_log10() +
    labs(x = "Barcode rank", y = "Total UMI count", title = samples$SampleName[i])
  print(p[[i]])

  e.out = emptyDrops(STAR[[i]]$RNA, lower = 1000, test.ambient = FALSE)
  is.cell = (e.out$FDR <= 0.01)
  w2kp = which(is.cell)

  STAR.ED[[i]]$RNA = STAR[[i]]$RNA[,w2kp]
  
  bcrank.ED = barcodeRanks(STAR.ED[[i]]$RNA, lower = 1000)
  
  knee_count <- paste0("Knee = ",table(bcrank.ED$total >= metadata(bcrank.ED)$knee)[2]," cells")
  inflection_count <- paste0("Inflection = ",table(bcrank.ED$total >= metadata(bcrank.ED)$inflection)[2]," cells")
  
  p_2[[i]] <- qplot(bcrank.ED$rank[!duplicated(bcrank.ED$rank)], bcrank.ED$total[!duplicated(bcrank.ED$rank)], geom = "line") +
    geom_hline(yintercept = metadata(bcrank.ED)$knee, color = "blue", linetype = 2) +
    geom_hline(yintercept = metadata(bcrank.ED)$inflection, color = "darkgreen", linetype = 2) +
    annotate("text", x = 10, y = 1.5 * c(metadata(bcrank.ED)$knee, metadata(bcrank.ED)$inflection),
           label = c(knee_count, inflection_count), color = c("blue", "darkgreen")) +
    scale_x_log10() +
    scale_y_log10() +
    labs(x = "Barcode rank", y = "Total UMI count", title = paste0(samples$SampleName[i]," ",ncol(STAR.ED[[i]]$RNA)," cells"))
  print(p_2[[i]])
}
dev.off()
```

```{r, eval = FALSE}
rm(bcrank, e.out, i, inflection_count, is.cell, knee_count, uniq, w2kp)
saveRDS(STAR.ED,"STAR_emptyDrops.rds")
```

```{r}
STAR <- readRDS("STAR_emptyDrops.rds")
```

```{r Pictures_EmptyDrop_filtered, echo=FALSE, fig.height=3, fig.width=4}
p_2 <- as.list(NULL)
pdf("knee_plot_EmptyDrops.pdf")
for(i in 1:nrow(samples)) {
  bcrank = barcodeRanks(STAR[[i]]$RNA, lower = 2000)
  uniq = !duplicated(bcrank$rank)
  
  knee_count <- paste0("Knee = ",table(bcrank$total >= metadata(bcrank)$knee)[2]," cells")
  inflection_count <- paste0("Inflection = ",table(bcrank$total >= metadata(bcrank)$inflection)[2]," cells")
  
  p_2[[i]] <- qplot(bcrank$rank[uniq], bcrank$total[uniq], geom = "line") +
    geom_hline(yintercept = metadata(bcrank)$knee, color = "blue", linetype = 2) +
    geom_hline(yintercept = metadata(bcrank)$inflection, color = "darkgreen", linetype = 2) +
    annotate("text", x = 10, y = 1.5 * c(metadata(bcrank)$knee, metadata(bcrank)$inflection),
           label = c(knee_count, inflection_count), color = c("blue", "darkgreen")) +
    scale_x_log10() +
    scale_y_log10() +
    labs(x = "Barcode rank", y = "Total UMI count", title = paste0(samples$SampleName[i]," ",ncol(STAR[[i]]$RNA)," cells"))
  print(p_2[[i]])
}
dev.off()
```

# Seurat, QC
```{r conversion, eval=FALSE, message=FALSE, warning=FALSE}
seurat<-as.list(NULL)
for(i in 1:nrow(samples)) {
  seurat[[as.vector(samples$SampleName[i])]] <- CreateSeuratObject(counts = STAR[[i]][["RNA"]], assay = "RNA")
  seurat[[i]]$Sample <- samples$SampleName[i]
  seurat[[i]]$Condition <- samples$Condition[i]
  seurat[[i]]$Genotype <- samples$Genotype[i]
  seurat[[i]]$Cultivation <- samples$Cultivation[i]
}
```

```{r eval=FALSE}
rm(STAR, p, p_2, i)
saveRDS(seurat,"seurat_emptyDrops.rds")
```

```{r eval=TRUE, include = FALSE}
seurat <- readRDS("seurat_emptyDrops.rds")
```

```{r}
# no probe for MALAT1 in fixed samples, mito is probably falsely mapped, and ribo is not there at all
seurat_nomalat <- as.list(NULL)
for(i in 1:nrow(samples)) {
  seurat_nomalat[[i]] <- seurat[[i]][!(rownames(seurat[[i]]) %in% c(grep("^RP[LS]",rownames(seurat[[i]]),value = TRUE),
                                                            grep("^MT-",rownames(seurat[[i]]),value = TRUE))), ]
  seurat_nomalat[[i]]$largest_count <- apply(seurat_nomalat[[i]]@assays$RNA@counts, 2, max)
  seurat_nomalat[[i]]$largest_index <- apply(seurat_nomalat[[i]]@assays$RNA@counts, 2, which.max)
  seurat_nomalat[[i]]$largest_gene <- rownames(seurat_nomalat[[i]])[seurat_nomalat[[i]]$largest_index]
  
  seurat_nomalat[[i]]$percent.Largest.Gene <- 100 * seurat_nomalat[[i]]$largest_count / seurat_nomalat[[i]]$nCount_RNA
  
  seurat[[i]]$largest_gene <- seurat_nomalat[[i]]$largest_gene
  seurat[[i]]$percent.Largest.gene <- seurat_nomalat[[i]]$percent.Largest.Gene
}
rm(seurat_nomalat)
gc()
```

```{r calculate_mt_ribo, eval = TRUE}
for(i in 1:nrow(samples)) {
  seurat[[i]][["percent.mt"]] <- PercentageFeatureSet(seurat[[i]], pattern = "^MT-")
}
```

```{r}
pdf("QCplots.pdf", width = 14, height = 7)
for(i in 1:nrow(samples)) {
print(gridExtra::grid.arrange(
  VlnPlot(seurat[[i]], features = c("nFeature_RNA"), pt.size = 0) + NoLegend() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    geom_hline(yintercept = 1200) + scale_y_log10(),
  VlnPlot(seurat[[i]], features = c("nCount_RNA"), pt.size = 0) + NoLegend() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    geom_hline(yintercept = 2500) + scale_y_log10(),
  VlnPlot(seurat[[i]], features = c("percent.mt"), pt.size = 0) + NoLegend() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    geom_hline(yintercept = 15),
  VlnPlot(seurat[[i]], features = c("percent.Largest.gene"), pt.size = 0) + scale_y_log10() + NoLegend() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)),
  ncol = 4, top = textGrob(names(seurat)[i]))
  )
}
dev.off()
```

```{r}
for(i in 1:nrow(samples)) {
 print(names(seurat)[i])
  largest_gene_list <- seurat[[i]]@meta.data %>%
  group_by(largest_gene) %>%
  count() %>%
  arrange(desc(n))
 print(largest_gene_list)
 write.csv(largest_gene_list,paste0(names(seurat)[i],"_largest_gene_list.csv"))
}
```

```{r eval=FALSE}
rm(i,largest_gene_list)
saveRDS(seurat,"seurat_QC.rds")
```

```{r eval = FALSE, include = FALSE}
seurat <- readRDS("seurat_QC.rds")
```

```{r Transfromation, eval=FALSE, message=FALSE, warning=FALSE}
all.genes <- rownames(seurat[[1]])
seurat_transformed <- as.list(NULL)
for(i in 1:nrow(samples)) {
  seurat_transformed[[i]] <- SCTransform(seurat[[i]],verbose = TRUE, return.only.var.genes = TRUE,assay = "RNA", n_genes = NULL, variable.features.n = NULL)
}
```

```{r eval=FALSE}
rm(i,p,all.genes)
saveRDS(seurat_transformed,"seurat_SCTransfrom.rds")
```

# Integration
```{r eval = FALSE, include = FALSE}
seurat_transformed <- readRDS("seurat_SCTransfrom.rds")
```

```{r Integration, eval=FALSE, message=FALSE, warning=FALSE}
gc()
all.genes <- rownames(seurat[[1]])
options(future.globals.maxSize = 100 * 1024^3)
seurat_features <- SelectIntegrationFeatures(object.list = seurat_transformed, nfeatures = 2000, fvf.nfeatures = 2000)
seurat_features <- seurat_features[substr(seurat_features,1,3) != "MT-"]
seurat_integrated <- PrepSCTIntegration(object.list = seurat_transformed, anchor.features = seurat_features, 
    verbose = TRUE)
seurat_Anchor <- FindIntegrationAnchors(object.list = seurat_integrated, normalization.method = "SCT", 
    anchor.features = seurat_features, verbose = TRUE, dims = 1:30, max.features = 500, n.trees = 50)
seurat_integrated <- IntegrateData(anchorset = seurat_Anchor, normalization.method = "SCT", 
    verbose = TRUE, dims = 1:30)
```

```{r eval=FALSE}
gc()
#saveRDS(seurat_Anchor,"seurat_Anchor.rds")
rm(seurat,seurat_Anchor,seurat_features,seurat_transformed,all.genes)
saveRDS(seurat_integrated,"seurat_integrated.rds")
```

```{r, eval = FALSE}
seurat_integrated <- readRDS("seurat_integrated.rds")
```

```{r PCA, eval = FALSE}
all.genes <- row.names(seurat_integrated)
seurat_integrated <- RunPCA(seurat_integrated, npcs = 50, verbose = FALSE, features = all.genes)
ElbowPlot(seurat_integrated, ndims = 50, reduction = "pca")
```

```{r UMAP, eval = FALSE}
seurat_integrated <- RunUMAP(seurat_integrated, reduction = "pca", dims = 1:20, verbose = TRUE)
seurat_integrated <- FindNeighbors(seurat_integrated, reduction = "pca", dims = 1:20, verbose = TRUE)
seurat_integrated <- FindClusters(seurat_integrated, resolution = 0.5, verbose = TRUE)
```

```{r eval=FALSE}
saveRDS(seurat_integrated,"seurat_integrated_UMAP.rds")
```

```{r eval = TRUE, include = FALSE}
seurat_integrated <- readRDS("seurat_integrated_UMAP.rds")
```

```{r pictures_plot, messages = FALSE, fig.width = 6, fig.height = 6}
dim.plot <- DimPlot(seurat_integrated, reduction = "umap", label = TRUE)
umap.rna <- FeaturePlot(seurat_integrated, features = c("nFeature_RNA","nCount_RNA"), min.cutoff = "q10", max.cutoff = "q90", ncol = 1, cols = c("gray","red"))
ggsave("UMAP_all.png", plot = dim.plot)
ggsave("umap_rna_QC.png", plot = umap.rna)
```

```{r}
#VlnPlot(seurat_integrated, features = c("nFeature_RNA","nCount_RNA","percent.mt","percent.rib"), ncol = 1)
pdf("QC_clusters.pdf")
  print(VlnPlot(seurat_integrated, features = "nFeature_RNA", ncol = 1, pt.size = 0) + geom_hline(yintercept=1500) + geom_hline(yintercept=2000, color = "red") + NoLegend())
  print(VlnPlot(seurat_integrated, features = "nCount_RNA", ncol = 1, pt.size = 0) + geom_hline(yintercept=2500) + ylim(2000,15000) + geom_hline(yintercept=3500, color = "red") + NoLegend())
  print(VlnPlot(seurat_integrated, features = "GFAP", ncol = 1, pt.size = 0) + geom_hline(yintercept=0) + NoLegend())
dev.off()
```

```{r, messages = FALSE, fig.width = 16}
pdf("UMAP_samples.pdf", )
  print(DimPlot(seurat_integrated, reduction = "umap", label = TRUE, split.by = "Sample"))
  print(DimPlot(seurat_integrated, reduction = "umap", label = FALSE, group.by = "Sample", pt.size = 1))
  dittoBarPlot(seurat_integrated, var = "seurat_clusters", group.by = "Condition")
dev.off()
```

```{r}
features_test <- c("NES","HSPB1","SYNM","CCDN2","SQSTM1","CRYAB",
                   "SOX9","NFIB",      # ASTRO transformation
                   "CD44","GJA1","S100B", #General ASTRO
                   "GFAP","VIM",                      # Reactive ASTRO
                   "EPCAM","CDH1","CLDN7",  #Maturated iPSC from fibroblats
                   "SLC17A6","ISL1",
                   "MAP2","SYP","DCX","STMN1","TH", #Neuronal
                   "CCND2","SOX4",
                   "MGP",
                   "VCAN",
                   "FABP7"
                 )
```

```{r, fig.height=5, fig.width = 10}
FeaturePlot(seurat_integrated, features = features_test, min.cutoff = "q10", max.cutoff = "q90", ncol = 6, cols = c("gray","red"))
DotPlot(seurat_integrated, features = features_test) + theme(axis.text.x=element_text(angle=90, hjust = 1, vjust = 0.5))
#FeaturePlot(seurat_integrated, features = "GFAP", min.cutoff = "q10", max.cutoff = "q90", cols = c("gray","red"))
```

# Markers
```{r Markers_Identification, eval = FALSE}
markers <- FindAllMarkers(seurat_integrated, min.pct = 0.8, only.pos = TRUE, verbose = TRUE, assay = "integrated")
```

```{r, eval = FALSE}
saveRDS(markers,"markers.rds")
write.csv(markers,"markers.csv")
```

```{r, eval = TRUE, include = FALSE}
markers <- readRDS("markers.rds")
```

```{r}
datatable(markers %>% group_by(cluster) %>% top_n(n = 15, wt = avg_log2FC))
```

# DoubletFinder
```{r eval = TRUE, include = FALSE}
seurat_integrated <- readRDS("seurat_integrated_UMAP.rds")
```

```{r, eval = FALSE}
seurat_SCT <- NULL
#Pre-process
for(i in 1:nrow(samples)) {
  seurat_SCT[[i]] <- seurat_integrated[,seurat_integrated$Sample == names(table(seurat_integrated$Sample))[i]]
  seurat_SCT[[i]] <- RunPCA(seurat_SCT[[i]], npcs = 20, verbose = FALSE)
  seurat_SCT[[i]] <- RunUMAP(seurat_SCT[[i]], dims = 1:20, verbose = FALSE)
  Idents(seurat_SCT[[i]]) <- as.matrix(Idents(seurat_integrated))[row.names(as.matrix(Idents(seurat_SCT[[i]]))),]
}
```

```{r eval=FALSE}
saveRDS(seurat_SCT,"seurat_SCT_UMAP.rds")
```

```{r read_SCT_after_UMAP, eval = TRUE, include = FALSE}
seurat_SCT <- readRDS("seurat_SCT_UMAP.rds")
```

```{r eval = FALSE, message=FALSE, warning=FALSE, include=TRUE}
#pK Identification (no ground-truth)
sweep.res.list <- as.list(NULL)
sweep.stats <- as.list(NULL)
bcmvn <- as.list(NULL)
bcmvn_pK_value <- as.vector(NULL)
p_bcmvn <- as.list(NULL)

for(i in 1:length(table(seurat_integrated$Sample))) {
  DefaultAssay(seurat_SCT[[1]]) <- "SCT"
  suppressWarnings(sweep.res.list[[i]] <- paramSweep_v3(seurat_SCT[[i]], PCs = 1:20, sct = TRUE, num.cores = 1))
  sweep.stats[[i]] <- summarizeSweep(sweep.res.list[[i]], GT = FALSE)
  bcmvn[[i]] <- find.pK(sweep.stats[[i]])
  
  bcmvn_BC_value <- max(bcmvn[[i]]$BCmetric)
  bcmvn_pK_value[i] <- as.numeric(as.vector(bcmvn[[i]][bcmvn[[i]]$BCmetric == bcmvn_BC_value,]$pK))
}
```

```{r eval = FALSE}
saveRDS(sweep.res.list,"DoubletFinder/sweep.res.list1.rds")
saveRDS(sweep.stats,"DoubletFinder/sweep.stats1.rds")
saveRDS(bcmvn_pK_value,"DoubletFinder/bcmvn_pK_value1.rds")
saveRDS(bcmvn,"DoubletFinder/bcmvn1.rds")
saveRDS(p_bcmvn,"DoubletFinder/p_bcmvn1.rds")
```

```{r read_pK_analysis, eval = TRUE, include = FALSE}
bcmvn_pK_value <- readRDS("bcmvn_pK_value1.rds")
bcmvn <- readRDS("bcmvn1.rds")
```

```{r print_pK_test, include = TRUE, eval = TRUE}
pdf("DoubletFinder/pK_values.pdf")
for(i in 1:length(table(seurat_integrated$Sample))) {
  p <- qplot(as.numeric(as.vector(bcmvn[[i]]$pK)), bcmvn[[i]]$BCmetric, geom = "line") +
    geom_vline(xintercept = bcmvn_pK_value[i], color = "red", linetype = 2) +
    annotate("text",
             x = if (bcmvn_pK_value[i] < 0.2) {
               bcmvn_pK_value[i] + 0.05
               } else {
               bcmvn_pK_value[i] - 0.05
               },
             y = 100,
             label = paste0("Optimal pK = ",bcmvn_pK_value[i]),
             color = "red") +
    labs(x = "pK", y = expression(BC[mvn]), title = names(table(seurat_integrated$Sample))[i])
  print(p)
}
dev.off()
```

```{r, eval = FALSE}
#Homotypic Doublet Proportion Estimate
annotations <- as.list(NULL)
homotypic.prop <- as.vector(NULL)
nExp_poi <- as.vector(NULL)
nExp_poi.adj <- as.vector(NULL)

for(i in 1:length(table(seurat_integrated$Sample))) {
  annotations[[i]] <- as.character(Idents(seurat_SCT[[i]]))
  homotypic.prop[i] <- modelHomotypic(annotations[[i]])
  nExp_poi[i] <- round(0.05*length(row.names(seurat_SCT[[i]]@meta.data)))  ## Assuming 5% doublet formation rate - estimated by 10x
  nExp_poi.adj[i] <- round(nExp_poi[i]*(1-homotypic.prop[i]))
}
```

```{r, eval = FALSE}
rm(annotations,homotypic.prop)
saveRDS(nExp_poi,"DoubletFinder/nExp_poi1.rds")
saveRDS(nExp_poi.adj,"DoubletFinder/nExp_poi_adj1.rds")
```

```{r read_nExp, eval = TRUE, include = FALSE}
nExp_poi <- readRDS("nExp_poi1.rds")
nExp_poi.adj <- readRDS("nExp_poi_adj1.rds")
```

```{r eval=FALSE, message=FALSE, warning=FALSE}
for(i in 1:length(table(seurat_integrated$Sample))) {
  suppressWarnings(seurat_SCT[[i]] <- doubletFinder_v3(seurat_SCT[[i]], PCs = 1:10, pN = 0.25, pK = bcmvn_pK_value[i], nExp = nExp_poi[i],
                                      reuse.pANN = FALSE, sct = TRUE))
  suppressWarnings(seurat_SCT[[i]] <- doubletFinder_v3(seurat_SCT[[i]], PCs = 1:10, pN = 0.25, pK = bcmvn_pK_value[i], nExp = nExp_poi.adj[i],
                                      reuse.pANN = paste0("pANN_0.25_",bcmvn_pK_value[i],"_",nExp_poi[i]), sct = TRUE))
}
```

```{r eval=FALSE}
saveRDS(seurat_SCT,"seurat_SCT_doubletFinder.rds")
```

```{r read_tables_with_doublets, eval=TRUE, include = FALSE}
seurat_SCT <- readRDS("seurat_SCT_doubletFinder.rds")
```

```{r, eval = FALSE}
seurat_dublets_all <- data.frame(NULL)

for(i in 1:length(table(seurat_integrated$Sample))) {
  seurat_dublets <- seurat_SCT[[i]]@meta.data[,colnames(seurat_SCT[[i]]@meta.data) == paste0("DF.classifications_0.25_",bcmvn_pK_value[i],"_",nExp_poi[i]) |
    colnames(seurat_SCT[[i]]@meta.data) == paste0("DF.classifications_0.25_",bcmvn_pK_value[i],"_",nExp_poi.adj[i])]
  colnames(seurat_dublets) <- c("Low Confidence","High Confidence")
  seurat_dublets_all <- rbind(seurat_dublets_all,seurat_dublets)
}

seurat_integrated$Doublet <- c(apply(seurat_dublets_all,1,
                                     function(x){
                                       if(x[1]==x[2])
                                         if(x[1]=="Singlet")
                                           "Singlet"
                                         else "Doublet - High Confidence"
                                       else "Doublet - Low Confidence"}))
```

```{r eval=FALSE}
gc()
rm(bcmvn,bcmvn_BC_value,bcmvn_pK_value,i,nExp_poi,nExp_poi.adj,p,p_bcmvn,seurat_dublets,seurat_dublets_all,seurat_SCT,sweep.res.list,sweep.stats)
saveRDS(seurat_integrated,"seurat_integrated_doubletFinder.rds")
```

```{r read_all_cells_RDS_with_doublets, eval=TRUE, include = FALSE}
seurat_integrated <- readRDS("seurat_integrated_doubletFinder.rds")
```

```{r, fig.width = 10, fig.height = 8}
pdf("DoubletFinder.pdf")
  DimPlot(seurat_integrated, reduction = "umap", label = TRUE)
  DimPlot(seurat_integrated, reduction = "umap", label = FALSE, group.by = "Doublet")
  FeaturePlot(seurat_integrated, features = c("percent.mt","nCount_RNA","nFeature_RNA"), min.cutoff = "q10", max.cutoff = "q90", ncol = 2, cols = c("gray","red"))
  DimPlot(seurat_integrated, reduction = "umap", label = FALSE, split.by = "Doublet")
  seurat_integrated$Clusters <- Idents(seurat_integrated)
  print(ggplot(data=seurat_integrated@meta.data, aes(Clusters)) +
    geom_bar(aes(fill=Doublet), position="fill") +
    theme(axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5)) +
    ylab('Proportion of cells'))

  print(ggplot(data=seurat_integrated@meta.data, aes(Sample)) +
    geom_bar(aes(fill=Doublet), position="fill") +
    theme(axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5)) +
    ylab('Proportion of cells'))
dev.off()
```

# Filtering for %MT-, nCount, nFeature, and Doublets, cell cycle scoring
```{r}
seurat_integrated <- readRDS("seurat_integrated_doubletFinder.rds")
seurat_filtered <- subset(seurat_integrated, (Doublet != "Doublet - High Confidence") & (percent.mt < 15))

VlnPlot(seurat_filtered, features = "percent.mt", pt.size = 0, group.by = "Sample")
VlnPlot(seurat_filtered, features = "nCount_RNA", pt.size = 0, group.by = "Sample")
VlnPlot(seurat_filtered, features = "nFeature_RNA", pt.size = 0, group.by = "Sample")
DimPlot(seurat_filtered, group.by = "Doublet")

saveRDS(seurat_filtered, "seurat_filtered.rds")

all.genes <- row.names(seurat_filtered)
seurat_filtered <- RunPCA(seurat_filtered, npcs = 50, verbose = FALSE, features = all.genes)
ElbowPlot(seurat_filtered, ndims = 50, reduction = "pca")

seurat_filtered <- RunUMAP(seurat_filtered, reduction = "pca", dims = 1:20, verbose = TRUE)
seurat_filtered <- FindNeighbors(seurat_filtered, reduction = "pca", dims = 1:20, verbose = TRUE)
seurat_filtered <- FindClusters(seurat_filtered, resolution = 0.5, verbose = TRUE)

umap.all <- DimPlot(seurat_filtered, label = T)
umap.split <- DimPlot(seurat_filtered, label = T, split.by = "Sample")
ggsave("UMAP_all_filtered.png", plot = umap.all, device = "png")
ggsave("UMAP_all_samples_filtered.png", plot = umap.split, device = "png", width = 10, height = 6)

seurat_filtered <- NormalizeData(seurat_filtered, normalization.method = "LogNormalize", scale.factor = 10000, assay = "RNA")
seurat_filtered <- ScaleData(seurat_filtered, features = all.genes, assay = "RNA")

markers <- FindAllMarkers(seurat_filtered, min.pct = 0.8, only.pos = TRUE, verbose = TRUE, assay = "RNA")

saveRDS(markers,"markers_filtered.rds")
write.csv(markers,"markers_filtered.csv")

datatable(markers %>% group_by(cluster) %>% top_n(n = 15, wt = avg_log2FC))

saveRDS(seurat_filtered, "seurat_filtered_markers.rds")
```

cell cycle scoring
```{r}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

seurat_filtered <- CellCycleScoring(seurat_filtered, s.features = s.genes, g2m.features = g2m.genes, set.ident = FALSE)

pdf("cell_cycle_score.pdf")
DimPlot(seurat_filtered, reduction = "umap", label = FALSE, split.by = "Phase")
dittoBarPlot(seurat_filtered, var = "Phase", group.by = "Sample", color.panel = c("red","green","blue"))
dittoBarPlot(seurat_filtered, var = "Phase", group.by = "seurat_clusters", color.panel = c("red","green","blue"))
dev.off()

saveRDS(seurat_filtered, "seurat_filtered_CCscore.rds")
```


# Subset to cortical and cerebral organoids:
Cerebral
```{r}
seurat_filtered <- readRDS("seurat_filtered.rds")
seurat_subset <- subset(seurat_filtered, Cultivation == "Cerebral") 

all.genes <- row.names(seurat_subset)
seurat_subset <- NormalizeData(seurat_subset, normalization.method = "LogNormalize", scale.factor = 10000, assay = "RNA")
seurat_subset <- ScaleData(seurat_subset, features = all.genes, assay = "RNA")
seurat_subset <- SCTransform(seurat_subset, assay = "RNA", verbose = TRUE, return.only.var.genes = TRUE, n_genes = NULL, variable.features.n = NULL)

seurat_subset <- RunPCA(seurat_subset, npcs = 50, verbose = FALSE, features = all.genes)
ElbowPlot(seurat_subset, ndims = 50, reduction = "pca")

seurat_subset <- RunUMAP(seurat_subset, reduction = "pca", dims = 1:20, verbose = TRUE)
seurat_subset <- FindNeighbors(seurat_subset, reduction = "pca", dims = 1:20, verbose = TRUE)
seurat_subset <- FindClusters(seurat_subset, resolution = 0.5, verbose = TRUE)

umap.all <- DimPlot(seurat_subset, label = T)
umap.split <- DimPlot(seurat_subset, label = T, split.by = "Sample")

markers <- FindAllMarkers(seurat_subset, min.pct = 0.8, only.pos = TRUE, verbose = TRUE, assay = "RNA", min.diff.pct = 0.2)

s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
seurat_subset <- CellCycleScoring(seurat_subset, s.features = s.genes, g2m.features = g2m.genes, set.ident = FALSE)

pdf("cerebral/cell_cycle_score_cerebral.pdf")
DimPlot(seurat_subset, reduction = "umap", label = FALSE, split.by = "Phase")
dittoBarPlot(seurat_subset, var = "Phase", group.by = "Sample", color.panel = c("red","green","blue"))
dittoBarPlot(seurat_subset, var = "Phase", group.by = "seurat_clusters", color.panel = c("red","green","blue"))
dev.off()

saveRDS(markers,"cerebral/markers_cerebral.rds")
write.csv(markers,"cerebral/markers_cerebral.csv")

ggsave("cerebral/UMAP_cerebral.png", plot = umap.all, device = "png")
ggsave("cerebral/UMAP_cerebral_samples.png", plot = umap.split, device = "png", width = 10, height = 6)
saveRDS(seurat_subset, "cerebral/seurat_subset_cerebral.rds")
```

Cortical
```{r}
seurat_filtered <- readRDS("seurat_filtered.rds")
seurat_subset <- subset(seurat_filtered, Cultivation == "Cortical") 

all.genes <- row.names(seurat_subset)
seurat_subset <- NormalizeData(seurat_subset, normalization.method = "LogNormalize", scale.factor = 10000, assay = "RNA")
seurat_subset <- ScaleData(seurat_subset, features = all.genes, assay = "RNA")
seurat_subset <- SCTransform(seurat_subset, assay = "RNA", verbose = TRUE, return.only.var.genes = TRUE, n_genes = NULL, variable.features.n = NULL)

seurat_subset <- RunPCA(seurat_subset, npcs = 50, verbose = FALSE, features = all.genes)
ElbowPlot(seurat_subset, ndims = 50, reduction = "pca")

seurat_subset <- RunUMAP(seurat_subset, reduction = "pca", dims = 1:20, verbose = TRUE)
seurat_subset <- FindNeighbors(seurat_subset, reduction = "pca", dims = 1:20, verbose = TRUE)
seurat_subset <- FindClusters(seurat_subset, resolution = 0.5, verbose = TRUE)

umap.all <- DimPlot(seurat_subset, label = T)
umap.split <- DimPlot(seurat_subset, label = T, split.by = "Sample")

markers <- FindAllMarkers(seurat_subset, min.pct = 0.8, only.pos = TRUE, verbose = TRUE, assay = "RNA", min.diff.pct = 0.2)

s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
seurat_subset <- CellCycleScoring(seurat_subset, s.features = s.genes, g2m.features = g2m.genes, set.ident = FALSE)

pdf("cortical/cell_cycle_score_cortical.pdf")
DimPlot(seurat_subset, reduction = "umap", label = FALSE, split.by = "Phase")
dittoBarPlot(seurat_subset, var = "Phase", group.by = "Sample", color.panel = c("red","green","blue"))
dittoBarPlot(seurat_subset, var = "Phase", group.by = "seurat_clusters", color.panel = c("red","green","blue"))
dev.off()

saveRDS(markers,"cortical/markers_cortical.rds")
write.csv(markers,"cortical/markers_cortical.csv")

ggsave("cortical/UMAP_cortical.png", plot = umap.all, device = "png")
ggsave("cortical/UMAP_cortical_samples.png", plot = umap.split, device = "png", width = 10, height = 6)
saveRDS(seurat_subset, "cortical/seurat_subset_cortical.rds")
```

Annotate neurons within neuronal clusters
```{r}
seurat_subset <- readRDS("cerebral/seurat_subset_cerebral.rds")
seurat_subset3 <- subset(seurat_subset, idents = c("5","6","7","8","9","10","11","17","18","19"))

all.genes <- row.names(seurat_subset3)
seurat_subset3 <- NormalizeData(seurat_subset3, normalization.method = "LogNormalize", scale.factor = 10000, assay = "RNA")
seurat_subset3 <- ScaleData(seurat_subset3, features = all.genes, assay = "RNA")
seurat_subset3 <- SCTransform(seurat_subset3, assay = "RNA", verbose = TRUE, return.only.var.genes = TRUE, n_genes = NULL, variable.features.n = NULL)

seurat_subset3 <- RunPCA(seurat_subset3, npcs = 50, verbose = FALSE, features = all.genes)
ElbowPlot(seurat_subset3, ndims = 50, reduction = "pca")

seurat_subset3 <- RunUMAP(seurat_subset3, reduction = "pca", dims = 1:20, verbose = TRUE)
seurat_subset3 <- FindNeighbors(seurat_subset3, reduction = "pca", dims = 1:20, verbose = TRUE)
seurat_subset3 <- FindClusters(seurat_subset3, resolution = 0.2, verbose = TRUE)

umap <- DimPlot(seurat_subset3, label = T)
ggsave("cerebral/UMAP_neuronal.png", umap, "png")

markers <- FindAllMarkers(seurat_subset3, min.pct = 0.8, only.pos = TRUE, verbose = TRUE, assay = "RNA", min.diff.pct = 0.2)
write.csv(markers, "cerebral/markers_neuronal.csv")

FeaturePlot(seurat_subset3, features = c("MKI67","TH","PAX6","SATB2", "EOMES", "HOPX"), min.cutoff = "q10", max.cutoff ="q90") 

saveRDS(seurat_subset3, "cerebral/seurat_neuronal_subset.rds")
```

# Cerebral annotation
```{r}
# mesenchymal-like/fibroblasts, clus0
clus0 <- c("COL1A1","DCN","LUM","DLK1","PDGFRA")
# ependymal/astrocytes
clus1 <- c("GFAP","CLU","S100B","SLC1A3","FOXJ1")
# satellite (muscle) cells
clus2 <- c("MYF5","NOTCH3","PAX7","FN1","MEGF10")
# pacreatic cells
clus3 <- c("REG3G","CLPS","CTRB1","CTRB2","CPA2")
# epithelial cells
clus4 <- c("EPCAM","ELF3","KRT8","NDRG1","KRT18")
# immature neurons
clus5 <- c("NEUROD2","NEUROD6","NRXN1","GRIA2","DISP3")
# peripheral neurons
clus6 <- c("TH","GATA2","GATA3","PHOX2A","PHOX2B","PRPH","ISL1")
# upper layer neurons
clus7 <- c("NEUROD2","NEUROD6","GRIA2","SATB2","FOXG1")
# VEGFA+ migrating? cells
clus8 <- c("VEGFA","SLC2A3","DDIT4","ENO2","PFKFB3")
# radial glia + oRG
clus9 <- c("SOX2","PAX6","GLI3","FABP7","TNC")
# intermediate progenitors
clus10 <- c("EOMES","NNAT","DCX")
# proliferating RG
clus11 <- c("TOP2A","NUSAP1","MKI67","SOX2","PAX6")
# proliferating satellite muscle cells
clus12 <- c("TOP2A","NUSAP1","MKI67","NOTCH3","FN1")
# mesothelial cells 
clus13 <- c("LRRN4","UPK3B","DCN","LUM","SERPINB9")
# mesenchymal-like cells/fibroblasts, ECM enriched
clus14 <- c("LUM","DLK1","DCN","COL15A1","IGFBP3")
# cardiac/skeletal muscle cells
clus15 <- c("MYOG","MYH3","TNNT2","TNNT3","ACTN2")
# choroid plexus
clus16 <- c("TRPM3","CA2","HTR2C","TTR","IGFBP7")
# immature neurons
clus17 <- c("NEUROD2","NEUROD6","FABP7","NNAT","STMN2")
# neuronal+mesenchymal-like markers -> doublets?
clus18 <- c("NEUROD6","GRIA2","DCX","PDGFRA","LUM")
# OPC
clus19 <- c("OLIG1","OLIG2")

markers.list <- c(clus0,clus1,clus2,clus3,clus4,
                  clus5,clus6,clus7,clus8,clus9,
                  clus10,clus11,clus12,clus13,clus14,
                  clus15,clus16,clus17,clus18,clus19)
markers.list <- unique(markers.list)
```

```{r}
seurat_subset_cerebral <- readRDS("cerebral/seurat_subset_cerebral.rds")
DimPlot(seurat_subset_cerebral, label = T) + NoLegend()
seurat_subset_cerebral <- RenameIdents(seurat_subset_cerebral, 
             "9" = "radial glia",
             "11" = "proliferating radial glia",
             "19" = "pre-OPCs",
             "10" = "intermediate progenitors",
             "5" = "immature neurons",
             "17" = "immature neurons NEUROD2+ FABP7+",
             "7" = "upper layer neurons",
             "6" = "peripheral neurons",
             "18" = "neuronal+mesenchymal-like cells",
             "1" = "astrocytes",
             "16" = "choroid plexus cells",
             "0" = "mesenchymal-like cells/fibroblasts",
             "14" = "mesenchymal-like cells/fibroblasts IGFBP3+ COL15A1+",
             "13" = "mesothelial cells",
             "2" = "satellite cells",
             "12" = "proliferating satellite cells",
             "15" = "cardiac/skeletal muscle cells",
             "4" = "epithelial cells",
             "3" = "pancreatic-like cells",
             "8" = "stressed VEGFA+ cells"
             )
             
seurat_subset_cerebral$clusters_annot <- seurat_subset_cerebral@active.ident
p <- DimPlot(seurat_subset_cerebral, label = F, pt.size = 1) + NoLegend() 
p <- LabelClusters(p, id = "ident",  fontface = "bold",size = 6,position = "nearest")

markers.list.new <- c("PAX6","FABP7","SOX2",
                      "OLIG1","EGFR","DLL3",
                      "EOMES","NNAT","NEUROD2","NEUROD6","NRXN1","GRIA2","SATB2","FOXG1",
                      "PHOX2B","PRPH","ISL1",
                      "FOXJ1","GFAP","S100B",
                      "TTR","TRPM3","CA2",
                      "COL1A1","DCN","LUM","PDGFRA","COL15A1","IGFBP3",
                      "MYF5","PAX7","FN1","MKI67","LRRN4","UPK3B","SERPINB9",
                      "MYOG","MYH3","TNNT2","EPCAM","ELF3","KRT8", 
                      "REG3G","CLPS","CTRB1",
                      "VEGFA","DDIT4","SLC2A3")

# levels(seurat_subset_cerebral) <- as.factor(c("mesenchymal-like/fibroblasts","mesenchymal-like/fibroblasts IGFBP3+ COL15A1+",
# "satellite cells","proliferating satellite cells","mesothelial cells","cardiac/skeletal muscle",
# "epithelial cells","pancreatic cells","radial glia","proliferating radial glia","OPC",
# "intermediate progenitors","immature neurons NEUROD2+ FABP7+","immature neurons","upper layer neurons","neuronal+mesenchymal-like","VEGFA+ cells","peripheral neurons","choroid plexus","ependymal/astrocytes"))
```

```{r}
seurat_subset_cerebral@active.assay <- "RNA"
goi.astro <- c("GFAP","AQP4")
goi.MG <- c("AIF1","TREM2")
goi.OPC <- c("OLIG1","OLIG2")
goi <- c(goi.astro,goi.MG,goi.OPC)

ggsave("cerebral/small_pops.png", plot = FeaturePlot(seurat_subset_cerebral, features = goi), device = "png", height = 10, width = 7)

nrow(seurat_subset_cerebral@meta.data[seurat_subset_cerebral$annot=="OPC" & seurat_subset_cerebral$Condition=="CER_cr",])
```

```{r}
seurat_subset_cerebral <- readRDS("cerebral/seurat_subset_cerebral.rds")
#palette <- alphabet(n = 20)
palette <- c("#CC7A88", "#E6B8BF", "#B33E52","#54990F", "#78B33E","#A3CC7A", "#CFE6B8","#0F8299", "#3E9FB3", "#7ABECC", "#B8DEE6","#653EB3", "#967ACC", "#C7B8E6","#B3823E", "#99600F",  "#CCAA7A", "#E6D2B8","#666666", "#999999") # palette <- stepped()

cluster_order <- match(levels(seurat_subset_cerebral), metaLevels("ident", seurat_subset_cerebral))
p.p <- dittoBarPlot(seurat_subset_cerebral, var = "ident" , group.by = "Condition", 
                    data.out = T, var.labels.reorder = cluster_order, color.panel = as.character(palette),main = "") 
p.p[["data"]]$percent100 <- p.p[["data"]]$percent * 100
data.p <- p.p[["data"]]

table(seurat_subset_cerebral$Condition)
```

Fig. 5A
```{r}
p1 <- DimPlot(seurat_subset_cerebral, pt.size = 0.5, cols = as.character(c("#CC7A88", "#E6B8BF", "#B33E52","#54990F", "#78B33E","#A3CC7A", "#CFE6B8","#0F8299", "#3E9FB3", "#7ABECC", "#B8DEE6","#653EB3", "#967ACC", "#C7B8E6","#B3823E", "#99600F",  "#CCAA7A", "#E6D2B8", "#999999")),label = F,
              cells = rownames(seurat_subset_cerebral@meta.data[seurat_subset_cerebral$Condition == "CER_cr",])) + 
  NoLegend() +
  xlim(-14,14) + ylim(-14,14) +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("CTRL")
p2 <- DimPlot(seurat_subset_cerebral, pt.size = 0.5,label = F, cols = as.character(palette),
              cells = rownames(seurat_subset_cerebral@meta.data[seurat_subset_cerebral$Condition == "CER_AxD",])) + NoLegend() +
  xlim(-14,14) + ylim(-14,14) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("AxD")

ggsave(filename = "230901_paper_figs/cerebral_UMAP.png", plot = grid.arrange(p1,p2,nrow = 2), device = "png",
       height = 10, width = 6)

data.p$label <- factor(data.p$label, levels = rev(levels(data.p$label)))
# barplot to complement the legend
p.p.bar2 <- ggplot(data = data.p, aes(x = count, y = label, fill = label)) + 
  geom_bar(stat = "identity", width = 0.9, position = "dodge2") +
  theme_void() +
  scale_x_reverse(limits = c(2600,0), expand = c(0.05,0), breaks = c(500,1500,2500), position = "top") +
  scale_y_discrete(position = "right") +
  scale_fill_manual(values = rev(as.character(palette))) +
  theme(axis.text.y = element_text(angle = 0, size = 11, hjust = 0, vjust = 0.3, 
                                   margin = margin(l=3,unit="pt")),
        axis.line.y = element_line(linewidth = 0.25, colour = "black"), 
        axis.ticks.y = element_line(linewidth = 0.25, colour = "black"),
        axis.line.x = element_line(linewidth = 0.25, colour = "black"),
        axis.ticks.x = element_line(linewidth = 0.25, colour = "black"),
        axis.ticks.length = unit(0.1, "cm"),
        axis.text.x = element_text(angle = 0, size = 6, hjust = 0.5, vjust = 1.5, colour = "black"),
        plot.margin = margin(r=0.7, b=0.5, t=0.5, l=0.5, unit="cm"),
        legend.position = "none") 
p.p.bar2 

ggsave("230901_paper_figs/Fig5A_legendbar.png", p.p.bar2, device = "png", bg = "white", height = 6, width = 5)
```

Fig. 5B
```{r}
seurat_subset_cerebral@active.assay <- "RNA"
goi <- c("PAX6","OLIG1","EOMES","FOXG1","ISL1","GFAP","TTR","DCN","TNNT2","EPCAM","CLPS","MKI67")
p <- FeaturePlot(seurat_subset_cerebral, features = goi, cols = c("#e0e0e0","#8c0052"), pt.size = 0.7, order = T, min.cutoff = "q10", max.cutoff = "q90", ncol = 3) + theme(plot.title = element_text(size = 14, face = "bold", hjust =  0.5)) & NoAxes() + NoLegend()

ggsave("230901_paper_figs/cerebral_goi2.png", p, height = 10, width = 9)

q <- FeaturePlot(seurat_subset_cerebral, "GFAP", split.by = "Condition", min.cutoff = "q10", max.cutoff = "q90", pt.size = 0.7, order = T, cols = c("#e0e0e0","#8c0052"), ) & NoAxes()
q
ggsave("230901_paper_figs/cerebral_GFAP.png", q, height = 4, width = 8)
```

Fig. C
```{r}
# RG DEA
DE_tests <- readRDS("cerebral/DEA_RG/CER_cr-CER_AxD.rds")
res <- DE_tests[[1]]

keyvals.colour <- ifelse(
  (res$avg_log2FC < -0.65 | res$avg_log2FC > 0.65) & res$p_val_adj < 0.05, "#8c0052",
         "#e0e0e0") 
names(keyvals.colour)[keyvals.colour == "#8c0052"] <- 'significant'
names(keyvals.colour)[keyvals.colour == "#e0e0e0"] <- 'insignificant'

labs <- c("CLPS","MOXD1","PCDH11X","REG3G","DNER","RTN1","HOPX","PTN","DLK1","COL3A1","CHGB","CRABP1","GAP43","PHOX2B","CNTN1","TTR","STMN2","HAND2","RGS5","CCND1")

p <- EnhancedVolcano(toptable = res,
                     lab = rownames(res),
                     selectLab = labs,
                     x = 'avg_log2FC',
                     y = 'p_val_adj',
                     axisLabSize = 8,
                     titleLabSize = 10,
                     title = "CTRL - AxD",
                     pCutoff = 0.05, 
                     colAlpha = 0.7,
                     colCustom = keyvals.colour,
                     subtitleLabSize = 9,
                     FCcutoff = 0.65, 
                     labSize = 3,
                     pointSize = 2.5,
                     drawConnectors = T,
                     maxoverlapsConnectors = Inf,
                     min.segment.length = 0.1,
                     arrowheads = FALSE,
                     boxedLabels = F,
                     cutoffLineType = "blank",
                     caption = "|log2FC| > 0.65 and p-adj < 0.05") +
  theme_light() +
  theme(plot.subtitle = element_blank(), 
        plot.caption = element_text(size = 10), 
        plot.title = element_text(hjust = 0.5, size = 12),
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        axis.line = element_line(colour = "black", linewidth = 0.7, linetype = "solid"), 
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        axis.ticks = element_line(linewidth = 0.7, colour = "black"),
        axis.ticks.length = unit(0.2, "cm"),
        legend.position = "none") +
  xlab(expression("Log"[2]*"FC")) +
  guides(color = guide_legend(override.aes = list(size = 4)))
p
ggsave(plot = p, "CER_volcanoRG.png", device = "png", width = 4.5, height = 3.5)

ego.results <- readRDS("cerebral/DEA_RG2/ORA_results.rds")
out <- merge_result(ego.results)
out@compareClusterResult <- out@compareClusterResult[order(out@compareClusterResult$p.adjust),]
out@compareClusterResult <- out@compareClusterResult[out@compareClusterResult$Count > 3,]
p <- dotplot(out, showCategory = 5, by = "GeneRatio", includeAll = F, label_format = 100) +
  theme_minimal() +
  theme(axis.text = element_text(size = 16, colour = "black"),
        axis.text.x = element_text(face = "bold"),
        axis.title.x = element_blank(),
        plot.title = element_text(size = 16, face = "bold", hjust = 8)) + 
  ggtitle("GO Terms Enrichment") +
  scale_color_gradient(low = "#8c0052",high = "#e0e0e0") 
p
ggsave(plot = p, filename = "CER_ORA_RG.png", device = "png", bg = "white",
       width = 8, height = 4)

# neurons DEA
DE_tests <- readRDS("cerebral_neuro_DEA/CER_cr-CER_AxD.rds")
res <- DE_tests[[1]]

keyvals.colour <- ifelse(
  (res$avg_log2FC < -0.65 | res$avg_log2FC > 0.65) & res$p_val_adj < 0.05, "#8c0052",
         "#e0e0e0") 
names(keyvals.colour)[keyvals.colour == "#8c0052"] <- 'significant'
names(keyvals.colour)[keyvals.colour == "#e0e0e0"] <- 'insignificant'

labs <- c("CLPS","ZNF454","PCDH11X","REG3G","PCDHA2","SPINK1","ZNF229","ZNF717","PCDHA3","PCDHGA8","PCDHGB6","ABR","CCSER1","MSRA","PKIB","ANK3","TTR","RUNX1T1","ISLR2","H1F0")

p <- EnhancedVolcano(toptable = res,
                     lab = rownames(res),
                     selectLab = labs,
                     x = 'avg_log2FC',
                     y = 'p_val_adj',
                     axisLabSize = 8,
                     titleLabSize = 10,
                     title = "CTRL - AxD",
                     pCutoff = 0.05, 
                     colAlpha = 0.7,
                     colCustom = keyvals.colour,
                     subtitleLabSize = 9,
                     FCcutoff = 0.65, 
                     labSize = 3,
                     pointSize = 2.5,
                     drawConnectors = T,
                     maxoverlapsConnectors = Inf,
                     min.segment.length = 0.1,
                     arrowheads = FALSE,
                     boxedLabels = F,
                     cutoffLineType = "blank",
                     caption = "|log2FC| > 0.65 and p-adj < 0.05") +
  theme_light() +
  theme(plot.subtitle = element_blank(), 
        plot.caption = element_text(size = 10), 
        plot.title = element_text(hjust = 0.5, size = 12),
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        axis.line = element_line(colour = "black", linewidth = 0.7, linetype = "solid"), 
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        axis.ticks = element_line(linewidth = 0.7, colour = "black"),
        axis.ticks.length = unit(0.2, "cm"),
        legend.position = "none") +
  xlab(expression("Log"[2]*"FC")) +
  guides(color = guide_legend(override.aes = list(size = 4)))
p
ggsave(plot = p, "CER_volcanoNEUR.png", device = "png", width = 4.5, height = 3.5)

ego.results <- readRDS("cerebral_neuro_DEA/ORA_results.rds")
out <- merge_result(ego.results)
out@compareClusterResult <- out@compareClusterResult[order(out@compareClusterResult$p.adjust),]
out@compareClusterResult <- out@compareClusterResult[out@compareClusterResult$Count > 3,]
p <- dotplot(out, showCategory = 5, by = "GeneRatio", includeAll = F, label_format = 40) +
  theme_minimal() +
  theme(axis.text = element_text(size = 16, colour = "black"),
        axis.text.x = element_text(face = "bold"),
        axis.title.x = element_blank(),
        plot.title = element_text(size = 16, face = "bold", hjust = 2)) + 
  ggtitle("GO Terms Enrichment") +
  scale_color_gradient(low = "#8c0052",high = "#e0e0e0") 
p
ggsave(plot = p, filename = "CER_ORA_NEUR.png", device = "png", bg = "white",
       width = 8, height = 4)
```

Supplementary Fig. 6C
```{r}
q.2 <- DotPlot(seurat_subset_cerebral,features = markers.list.new) +
  theme(axis.text.x = element_text(angle = 40, vjust = 0.9, hjust = 0.9)) +
  scale_color_gradient(high = "#8c0052",low = "#e0e0e0")
ggsave("Supp_Fig6C_CERmarkers.png", plot = q.2, device = "png", bg = "white", width = 20, height = 6)
```

Supplementary Fig. 6E
```{r}
seurat.object <- seurat_subset_cerebral
gene.set <- list("Glycolysis GO:0006096 Score" = as.vector(read.table("stressed_cells/glycolysisGO.txt", header = F)[,1]),
                 "ER Stress GO:0034976 Score" = as.vector(read.table("stressed_cells/ERstressGO.txt", header = F)[,1]))

stress_score <- function(seurat.object, genes, title){
  seurat.scored <- AddModuleScore(seurat.object, features = list(genes), assay = "RNA", name = "reference_score_")
  p <- FeaturePlot(seurat.scored, features = "reference_score_1", order = T,
                   min.cutoff = "q70", max.cutoff = "q90", reduction = "umap") + 
    ggtitle(title) + scale_color_gradient(high = "#8c0052",low = "#e0e0e0") 
  print(p)
  
  return(p)
}
  
stress_plots <- lapply(seq_along(gene.set), function(x) stress_score(seurat.object, gene.set[[x]], title = names(gene.set)[x]))  
ggsave(filename = "Supp_Fig6E_CER_stressed_GOterms.png", plot = grid.arrange(grobs = stress_plots, nrow = 1), device = "png", width = 11, height = 5)
```

# Cortical annotation
Cortical control clustering and pseudotime
```{r}
seurat_subset_cortical
seurat_cortical_CTRLneurons <- subset(seurat_subset_cortical, Condition == "Ctx_cr")

all.genes <- row.names(seurat_cortical_CTRLneurons)
seurat_cortical_CTRLneurons <- NormalizeData(seurat_cortical_CTRLneurons, normalization.method = "LogNormalize", scale.factor = 10000, assay = "RNA")
seurat_cortical_CTRLneurons <- ScaleData(seurat_cortical_CTRLneurons, features = all.genes, assay = "RNA")
seurat_cortical_CTRLneurons <- SCTransform(seurat_cortical_CTRLneurons, assay = "RNA", verbose = TRUE, return.only.var.genes = TRUE, n_genes = NULL, variable.features.n = NULL)

seurat_cortical_CTRLneurons <- RunPCA(seurat_cortical_CTRLneurons, npcs = 50, verbose = FALSE, features = all.genes)
ElbowPlot(seurat_cortical_CTRLneurons, ndims = 50, reduction = "pca")

seurat_cortical_CTRLneurons <- RunUMAP(seurat_cortical_CTRLneurons, reduction = "pca", dims = 1:20, verbose = TRUE)
seurat_cortical_CTRLneurons <- FindNeighbors(seurat_cortical_CTRLneurons, reduction = "pca", dims = 1:20, verbose = TRUE)
seurat_cortical_CTRLneurons <- FindClusters(seurat_cortical_CTRLneurons, resolution = 0.3, verbose = TRUE)

umap <- DimPlot(seurat_cortical_CTRLneurons, label = T)
ggsave("cortical/UMAP_CTRLneuronal.png", umap, "png")

markers <- FindAllMarkers(seurat_cortical_CTRLneurons, min.pct = 0.8, only.pos = TRUE, verbose = TRUE, assay = "RNA")
write.csv(markers, "cortical/markers_CTRLneuronal.csv")

FeaturePlot(seurat_cortical_CTRLneurons, features = c("MKI67","LHX6","PAX6","GAP43", "HOPX"), min.cutoff = "q10", max.cutoff ="q90") 

saveRDS(seurat_cortical_CTRLneurons, "cortical/seurat_cortical_CTRLneurons.rds")
```

Cortical annotation, not integrated
```{r}
seurat_subset_cortical <- readRDS("cortical/seurat_subset_cortical.rds")
DimPlot(seurat_subset_cortical, label = T) + NoLegend()
seurat_subset_cortical <- RenameIdents(seurat_subset_cortical, 
             "0" = "neurons 3",
             "1" = "neurons 1",
             "2" = "AxD neurons",
             "3" = "AxD progenitors",
             "4" = "mesenchymal-like",
             "5" = "cardiac/skeletal muscle",
             "6" = "neurons 2",
             "7" = "proliferating mesenchymal-like",
             "8" = "cycling progenitors",
             "9" = "proliferating AxD progenitors",
             "10" = "transitioning cells",
             "11" = "oRG",
             "12" = "cortical interneurons",
             "13" = "pyramidal neurons",
             "14" = "fibroblasts",
             "15" = "SPARCL1+ AxD cluster",
             "16" = "CRABP1+ VIM+ AxD cluster")

palette <- alphabet()
palette <- palette[c(1,2,3,7,8,9,10,11,12,13,15,16,17,18,19,20,22)]
p <- DimPlot(seurat_subset_cortical, label = F, pt.size = 1, cols = as.character(palette)) + NoLegend() 
p <- LabelClusters(p, id = "ident",  fontface = "bold",size = 6)
ggsave("cortical/UMAP_cortical_annot2.png",plot = p, device = "png", width = 12, height = 12)

p.2 <- DimPlot(seurat_subset_cortical, group.by = "Condition", label = F)
ggsave("cortical/UMAP_sample.png", plot = p.2, width = 7, height = 5)

table(seurat_subset_cortical$Condition)

markers.list.new <- c("DCN","COL1A1","MKI67","FN1","LUM","PDGFRA",
                      "TIMP1","HSPA1B","CRYAB",
                      "MYH3","MYOG",
                      "HES5","PAX6","SOX2",
                      "HES6","ASCL1","GSX2","FABP7",
                      "MOXD1","FAM107A","TNC",
                      "FOXG1","DLX2","NNAT","NRXN1","DCX","GAD1","GDAP1","STMN2",
                      "BCL11B","GRIA2",
                      "NXPH1","LHX6",
                      "NEUROD2","NEUROD6","TBR1",
                      "GAP43","GRIA2","STMN2",
                      "CLU","ATP1A2","SPARCL1",
                      "CRABP1","NR2F1")
markers.list.new <- unique(markers.list.new)

levels(seurat_subset_cortical) <- as.factor(c("mesenchymal-like","proliferating mesenchymal-like","fibroblasts","transitioning cells","cardiac/skeletal muscle",
"AxD progenitors","proliferating AxD progenitors","cycling progenitors","oRG",
"neurons 1","neurons 2","neurons 3",
"cortical interneurons","pyramidal neurons","AxD neurons",
"SPARCL1+ AxD cluster","CRABP1+ VIM+ AxD cluster"))
q.2 <- DotPlot(seurat_subset_cortical,features = markers.list.new) +
  theme(axis.text.x = element_text(angle = 40, vjust = 0.9, hjust = 0.9))
q.2
ggsave("cortical/markers_dotplot2.png", plot = q.2, device = "png", bg = "white", width = 20, height = 6)
```

Pollen 2015 radial glia
```{r}
dir <- dir("reference_datasets/Pollen2015/", pattern = ".txt")
dir.link <- "reference_datasets/Pollen2015/"

pdf("cortical/neurons_Pollen2015_RG.pdf", width = 15, height = 15)
for(i in 1:length(dir)){
  gene.set <- as.vector(read.table(paste0(dir.link,dir[i]), header = T)[,1])
  seurat.scored <- AddModuleScore(seurat_cortical_CTRLneurons, features = list(gene.set), assay = "RNA", name = "reference_score_")
  p <- FeaturePlot(seurat.scored, features = "reference_score_1", 
                   min.cutoff = "q10", max.cutoff = "q90", reduction = "umap") +
    ggtitle(dir[i])
  print(p)
}
dev.off()

pdf("cortical/Pollen2015_RG.pdf", width = 15, height = 15)
for(i in 1:length(dir)){
  gene.set <- as.vector(read.table(paste0(dir.link,dir[i]), header = T)[,1])
  seurat.scored <- AddModuleScore(seurat_subset_cortical, features = list(gene.set), assay = "RNA", name = "reference_score_")
  p <- FeaturePlot(seurat.scored, features = "reference_score_1", 
                   min.cutoff = "q10", max.cutoff = "q90", reduction = "umap") +
    ggtitle(dir[i])
  print(p)
}
dev.off()
```

Polioudakis2019's annotation
```{r}
dir <- dir("reference_datasets/Polioudakis2019/all/", pattern = ".txt")
dir.link <- "reference_datasets/Polioudakis2019/all/"

pdf("cortical/Polioudakis_all_neurons_overlap.pdf", width = 10, height = 10)
for(i in 1:length(dir)){
  gene.set <- as.vector(read.table(paste0(dir.link,dir[i]), header = T)[,1])
  seurat.scored <- AddModuleScore(seurat_cortical_CTRLneurons, features = list(gene.set), assay = "RNA", name = "reference_score_")
  p <- FeaturePlot(seurat.scored, features = "reference_score_1", 
                   min.cutoff = "q10", max.cutoff = "q90", reduction = "umap") +
    ggtitle(dir[i])
  print(p)
}
dev.off()
```

# Subset for DEA
```{r}
seurat_subset <- readRDS("cortical/seurat_subset_cortical.rds")

#cerebral
seurat_subset <- seurat_subset_cerebral
seurat_subset1 <- subset(seurat_subset, idents = c("10","17","5","7","18")) # neuro
seurat_subset1a <- subset(seurat_subset, subset = seurat_clusters %in% c("10","17","5","7")) # neuro new
seurat_subset2 <- subset(seurat_subset, idents = c("9","11","8","19")) #RG
seurat_subset3 <- subset(seurat_subset, idents = c("9","11")) #RG without VEGFA+ and OPC

#cortical
seurat_subset1 <- subset(seurat_subset, idents = c("1","6","0","12","13","2")) # neuro
seurat_subset2 <- subset(seurat_subset, idents = c("8","11","3","16","9","15")) #RG

p <- DimPlot(seurat_subset, cells.highlight = list(as.character(colnames(seurat_subset1)), as.character(colnames(seurat_subset2))), cols.highlight = c("#74C2BB","#FA8072")) + NoLegend()
ggsave("cortical/DEA_selection.png", plot=p, device = "png", width = 7, height = 7)

seurat_for_DEA <- function(so){
  all.genes <- row.names(so)
  so <- NormalizeData(so, normalization.method = "LogNormalize", scale.factor = 10000, assay = "RNA")
  so <- ScaleData(so, features = all.genes, assay = "RNA")
  so <- SCTransform(so, assay = "RNA", verbose = TRUE, return.only.var.genes = TRUE, n_genes = NULL, variable.features.n = NULL)
  
  so <- RunPCA(so, npcs = 50, verbose = FALSE, features = all.genes)
  #ElbowPlot(so, ndims = 50, reduction = "pca")
  
  so <- RunUMAP(so, reduction = "pca", dims = 1:15, verbose = TRUE)
  so <- FindNeighbors(so, reduction = "pca", dims = 1:15, verbose = TRUE)
  so <- FindClusters(so, resolution = 0.2, verbose = TRUE)
  
  return(so)
}

saveRDS(seurat_subset3, "cortical/seurat_neuro_subset_forDEA.rds")
saveRDS(seurat_subset3, "cortical/seurat_RG_subset_forDEA.rds")

seurat.DEA <- seurat_for_DEA(seurat_subset1a)
saveRDS(seurat.DEA, "230901_paper_figs/cerebral_neuro_DEA/seurat_neuro_forDEA.rds")
```

DEA
```{r}
source("DEA.R")
```

# Cortical integration
```{r}
seurat.object <- readRDS("cortical/seurat_subset_cortical.rds")

seurat.list <- SplitObject(seurat.object, split.by = "Condition")
seurat.list <- lapply(X = seurat.list, FUN = SCTransform)

options(future.globals.maxSize = 100 * 1024^3)
seurat_features <- SelectIntegrationFeatures(object.list = seurat.list, nfeatures = 2000, fvf.nfeatures = 2000)
seurat_features <- seurat_features[substr(seurat_features,1,3) != "MT-"]
seurat.list <- PrepSCTIntegration(object.list = seurat.list, anchor.features = seurat_features, 
    verbose = TRUE)
seurat_Anchor <- FindIntegrationAnchors(object.list = seurat.list, normalization.method = "SCT", 
    anchor.features = seurat_features, verbose = TRUE, dims = 1:30, max.features = 500, n.trees = 50)
seurat_integrated <- IntegrateData(anchorset = seurat_Anchor, normalization.method = "SCT", 
    verbose = TRUE, dims = 1:30)

rm(seurat.object,seurat_Anchor,seurat_features,seurat.list)

all.genes <- row.names(seurat_integrated)
seurat_integrated <- RunPCA(seurat_integrated, npcs = 50, verbose = FALSE, features = all.genes)
ElbowPlot(seurat_integrated, ndims = 50, reduction = "pca")
seurat_integrated <- RunUMAP(seurat_integrated, reduction = "pca", dims = 1:10, verbose = TRUE)
seurat_integrated <- FindNeighbors(seurat_integrated, reduction = "pca", dims = 1:10, verbose = TRUE)
seurat_integrated <- FindClusters(seurat_integrated, resolution = 0.3, verbose = TRUE)

a <- DimPlot(seurat_integrated, label = T)
b <- DimPlot(seurat_integrated, group.by = "Condition")
a|b
ggsave("cortical_integrated/UMAP_cortical.png", plot = a, device = "png", width = 7, height = 7)
ggsave("cortical_integrated/UMAP_cortical_sample.png", plot = b, device = "png", width = 5, height = 4)

seurat_integrated@active.assay <- "RNA"
all.genes <- row.names(seurat_integrated)
seurat_integrated <- NormalizeData(seurat_integrated, normalization.method = "LogNormalize", scale.factor = 10000, assay = "RNA")
seurat_integrated <- ScaleData(seurat_integrated, features = all.genes, assay = "RNA")

saveRDS(seurat_integrated,"cortical_integrated/seurat_cortical_integrated.rds")

table(seurat_integrated$Condition)
```

```{r}
markers <- FindAllMarkers(seurat_integrated, min.pct = 0.5, only.pos = TRUE, verbose = TRUE, assay = "RNA", min.diff.pct = 0.2)
write.csv(markers, "cortical_integrated/markers_integrated.csv")

markers.list <- unique(c("HES5","PAX6","SOX2",
                         "MKI67","TNC","HOPX","MOXD1",
                         "FABP7","AQP4","SOX9",
                         "CRABP1",
                         "ASCL1","GSX2","SMOC1",
                         "ELAVL4","DCX","STMN2",
                         "SYT4","GRIA2","GAP43",
                         "GAD1","GRIA1","PCDHB7",
                         "DCN","COL1A1","FN1","MKI67",
                         "ACTC1","TNNT2","MYH8",
                         "VEGFA","GADD45A","P4HA1"
                         ))

seurat_integrated <- RenameIdents(seurat_integrated,
                                  "2" = "radial glia",
                                  "8" = "proliferating radial glia",
                                  "9" = "oRG/astrocytes",
                                  "11" = "CRABP1+ cells",
                                  "10" = "pre-OPCs",
                                  "4" = "intermediate progenitors",
                                  "0" = "neurons 1",
                                  "1" = "neurons 2",
                                  "3" = "mesenchymal-like cells",
                                  "7" = "proliferating mesenchymal-like cells",
                                  "5" = "muscle cells",
                                  "6" = "stressed VEGFA+ cells")

q.2 <- DotPlot(seurat_integrated,features = markers.list) +
  theme(axis.text.x = element_text(angle = 40, vjust = 0.9, hjust = 0.9)) +
  scale_color_gradient(high = "#8c0052",low = "#e0e0e0")
ggsave("Supp_Fig6D_CTXmarkers.png", plot = q.2, device = "png", bg = "white", width = 14, height = 4)

seurat_integrated$Condition <- factor(seurat_integrated$Condition, levels = c("Ctx_cr", "Ctx_AxD"))

cluster_order <- match(levels(seurat_integrated), metaLevels("ident", seurat_integrated))
bar.p <- dittoBarPlot(seurat_integrated, var = "ident" , group.by = "Condition", 
                    data.out = T, var.labels.reorder = cluster_order, color.panel = as.character(palette),main = "") 
bar.p[["data"]]$percent100 <- bar.p[["data"]]$percent * 100
data.p <- bar.p[["data"]]
```

Fig. 6A
```{r}
palette <- c("#CC7A88", "#E6B8BF", "#B33E52", "#7ABECC", "#3E9FB3","#54990F", "#78B33E","#A3CC7A", "#653EB3", "#967ACC", "#C7B8E6", "#999999")
# latest figure version
p1 <- DimPlot(seurat_integrated, pt.size = 0.5, cols = as.character(palette),label = F,
              cells = rownames(seurat_integrated@meta.data[seurat_integrated$Condition == "Ctx_cr",])) + 
  NoLegend() +
  xlim(-12,12) + ylim(-12,12) +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("CTRL")
p2 <- DimPlot(seurat_integrated, pt.size = 0.5,label = F, cols = as.character(palette),
              cells = rownames(seurat_integrated@meta.data[seurat_integrated$Condition == "Ctx_AxD",])) + 
  NoLegend() +
  xlim(-12,12) + ylim(-12,12) +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("AxD")

ggsave(filename = "230901_paper_figs/cortical_UMAP.png", plot = grid.arrange(p1,p2,nrow = 2), device = "png",
       height = 10, width = 6)

data.p$label <- factor(data.p$label, levels = rev(levels(data.p$label)))
# barplot to complement the legend
p.p.bar2 <- ggplot(data = data.p, aes(x = count, y = label, fill = label)) + 
  geom_bar(stat = "identity", width = 0.9, position = "dodge2") +
  theme_void() +
  scale_x_reverse(limits = c(1800,0), expand = c(0.05,0), breaks = c(500,1000,1500), position = "top") +
  scale_y_discrete(position = "right") +
  scale_fill_manual(values = rev(as.character(palette))) +
  theme(axis.text.y = element_text(angle = 0, size = 11, hjust = 0, vjust = 0.3, 
                                   margin = margin(l=3,unit="pt")),
        axis.line.y = element_line(linewidth = 0.25, colour = "black"), 
        axis.ticks.y = element_line(linewidth = 0.25, colour = "black"),
        axis.line.x = element_line(linewidth = 0.25, colour = "black"),
        axis.ticks.x = element_line(linewidth = 0.25, colour = "black"),
        axis.ticks.length = unit(0.1, "cm"),
        axis.text.x = element_text(angle = 0, size = 6, hjust = 0.5, vjust = 1.5, colour = "black"),
        plot.margin = margin(r=0.7, b=0.5, t=0.5, l=0.5, unit="cm"),
        legend.position = "none") 
p.p.bar2 
ggsave("230901_paper_figs/Fig6A_legendbar.png", p.p.bar2, device = "png", bg = "white", height = 5, width = 4)
```

Subset for DEA
```{r}
seurat.object <- readRDS("cortical_integrated/seurat_cortical_integrated.rds")
seurat.object@active.ident <- seurat.object$integrated_snn_res.0.3
seurat.subset <- subset(seurat.object, idents = c("0","1","2","4","8","9","10","11")) # all except mesench. and muscle cells and VEGFA+
seurat.subset <- subset(seurat.object, idents = c("9")) #oRG

p <- DimPlot(seurat.object, cells.highlight = list(as.character(colnames(seurat.subset))), cols.highlight = "#74C2BB") + NoLegend()
ggsave("cortical_integrated/DEA_selection_oRG.png", plot=p, device = "png", width = 7, height = 7)

all.genes <- row.names(seurat.subset)
seurat.subset <- NormalizeData(seurat.subset, normalization.method = "LogNormalize", scale.factor = 10000, assay = "RNA")
seurat.subset <- ScaleData(seurat.subset, features = all.genes, assay = "RNA")
seurat.subset <- SCTransform(seurat.subset, assay = "RNA", verbose = TRUE, return.only.var.genes = TRUE, n_genes = NULL, variable.features.n = NULL)

seurat.subset <- RunPCA(seurat.subset, npcs = 50, verbose = FALSE, features = all.genes)
ElbowPlot(seurat.subset, ndims = 50, reduction = "pca")

seurat.subset <- RunUMAP(seurat.subset, reduction = "pca", dims = 1:15, verbose = TRUE) # neural 15, oRG 10
seurat.subset <- FindNeighbors(seurat.subset, reduction = "pca", dims = 1:15, verbose = TRUE)
seurat.subset <- FindClusters(seurat.subset, resolution = 0.2, verbose = TRUE)

saveRDS(seurat.subset, "cortical_integrated/seurat_neural_forDEA.rds")
saveRDS(seurat.subset, "cortical_integrated/seurat_oRG_forDEA.rds")
```

DEA
```{r}
source("DEA.R")
```

Fig. 6B
```{r}
seurat_cortical_integrated <- readRDS("seurat_cortical_integrated.rds")

# higlight markers
seurat_cortical_integrated@active.assay <- "RNA"
goi <- c("PAX6","GFAP","FOXG1","DCX","DCN","TNNT2","MKI67","VEGFA")
p <- FeaturePlot(seurat_cortical_integrated, features = goi, cols = c("#e0e0e0","#8c0052"), pt.size = 0.7, order = T, min.cutoff = "q10", max.cutoff = "q90", ncol = 4) + theme(plot.title = element_text(size = 14, face = "bold", hjust =  0.5)) & NoAxes() + NoLegend()

ggsave("cortical_goi.png", p, height = 6, width = 11)
```

Fig. 6C
```{r}
seurat_cortical_integrated@active.ident <- as.factor(seurat_cortical_integrated$Condition)
seurat_cortical_integrated <- RenameIdents(seurat_cortical_integrated,
                                           "Ctx_cr" = "CTRL",
                                           "Ctx_AxD" = "AxD")
seurat_cortical_integrated$Condition <- seurat_cortical_integrated@active.ident
seurat_cortical_integrated@active.ident <- as.factor(seurat_cortical_integrated$clusters_annot)
p <- VlnPlot(seurat_cortical_integrated, "FOXG1", split.by = "Condition",
             idents = c("intermediate progenitors","neurons 1","neurons 2"), 
             cols = c("#639DD1","#e06666")) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
        axis.title.x = element_blank(),
        legend.position = c(0.85,0.9)) +
  scale_x_discrete(labels = c("intermediate \nprogenitors", "neurons 1", "neurons 2"))
p
ggsave(plot = p, filename = "Fig_6C_cortical_FOXG1.png", device = "png",
       width = 6, height = 4)
```

Fig. 6D
```{r}
# RG DEA
DE_tests <- readRDS("cortical_integrated/DEA_oRG/Ctx_cr-Ctx_AxD.rds")
res <- DE_tests[[1]]

keyvals.colour <- ifelse(
  (res$avg_log2FC < -0.65 | res$avg_log2FC > 0.65) & res$p_val_adj < 0.05, "#8c0052",
         "#e0e0e0") 
names(keyvals.colour)[keyvals.colour == "#8c0052"] <- 'significant'
names(keyvals.colour)[keyvals.colour == "#e0e0e0"] <- 'insignificant'

labs <- c("NFIB","SOX9","NNAT","NRXN1","AQP4","NCAN","FABP7","FAM171B","BCAN","FOXG1","MOXD1","C1orf61",
          "SLIT1","EFNB3","RSPO3","CHL1","ID3","CRABP2","EPHA7","FN1","NPTX2","SPARCL1")

p <- EnhancedVolcano(toptable = res,
                     lab = rownames(res),
                     selectLab = labs,
                     x = 'avg_log2FC',
                     y = 'p_val_adj',
                     axisLabSize = 8,
                     titleLabSize = 10,
                     title = "CTRL - AxD",
                     pCutoff = 0.05, 
                     colAlpha = 0.7,
                     colCustom = keyvals.colour,
                     subtitleLabSize = 9,
                     FCcutoff = 0.65, 
                     labSize = 3,
                     pointSize = 2.5,
                     drawConnectors = T,
                     maxoverlapsConnectors = Inf,
                     min.segment.length = 0.1,
                     arrowheads = FALSE,
                     boxedLabels = F,
                     cutoffLineType = "blank",
                     caption = "|log2FC| > 0.65 and p-adj < 0.05"
                     ) +
  theme_light() +
  theme(plot.subtitle = element_blank(), 
        plot.caption = element_text(size = 10), 
        plot.title = element_text(hjust = 0.5, size = 12),
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        axis.line = element_line(colour = "black", linewidth = 0.7, linetype = "solid"), 
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        axis.ticks = element_line(linewidth = 0.7, colour = "black"),
        axis.ticks.length = unit(0.2, "cm"),
        legend.position = "none") +
  xlab(expression("Log"[2]*"FC")) +
  guides(color = guide_legend(override.aes = list(size = 4)))
p
ggsave(plot = p, "Fig_6D_CTX_volcanoRG.png", device = "png", width = 4.5, height = 3.5)

ego.results <- readRDS("cortical_integrated/DEA_oRG/ORA_results.rds")
out <- merge_result(ego.results)
out@compareClusterResult <- out@compareClusterResult[order(out@compareClusterResult$p.adjust),]
out@compareClusterResult <- out@compareClusterResult[out@compareClusterResult$Count > 3,]
p <- dotplot(out, showCategory = 5, by = "GeneRatio", includeAll = F, label_format = 100) +
  theme_minimal() +
  theme(axis.text = element_text(size = 16, colour = "black"),
        axis.text.x = element_text(face = "bold"),
        axis.title.x = element_blank(),
        plot.title = element_text(size = 16, face = "bold", hjust = 8)) + 
  ggtitle("GO Terms Enrichment") +
  scale_color_gradient(low = "#8c0052",high = "#e0e0e0") 
  
p

ggsave(plot = p, filename = "Fig_6D_CTX_ORA_oRG.png", device = "png", bg = "white",
       width = 8, height = 4)
```


# Supplement
Supp. Fig. 6A
```{r}
seurat_integrated <- readRDS("seurat_integrated.rds")
seurat_integrated <- SetIdent(seurat_integrated, value = seurat_integrated$Sample)
seurat_integrated <- RenameIdents(seurat_integrated, "CER_cr" = "UNG CTRL", 
                                                     "CER_AxD" = "UNG AxD", 
                                                     "Ctx_cr" = "CTX CTRL", 
                                                     "Ctx_AxD" = "CTX AxD")
vlnplotCustom <- function(features, intercept=NULL){
  return(VlnPlot(seurat_integrated, features = features, pt.size = 0) +
           geom_hline(yintercept = intercept) +
           NoLegend() +
           theme(axis.title.x = element_blank(),
                   plot.title = element_text(size = 12)))
}
p1<-vlnplotCustom(features = "nCount_RNA") + theme(axis.text.x = element_blank())
p2<-vlnplotCustom(features = "nFeature_RNA") + theme(axis.text.x = element_blank())
p3<-vlnplotCustom(features = "percent.mt", intercept = 10)
ggsave("Supp_Fig6A_VlnQC.png", plot = egg::ggarrange(p1,p2,p3, nrow = 3), device = "png", width = 5, height = 7)
```

Supp. Fig. 6B
```{r}
seurat_filtered_markers <- readRDS("seurat_filtered_markers.rds")
seurat_filtered_markers <- SetIdent(seurat_filtered_markers, value = seurat_filtered_markers$Sample)
seurat_filtered_markers <- RenameIdents(seurat_filtered_markers, "CER_cr" = "UNG CTRL", 
                                                     "CER_AxD" = "UNG AxD", 
                                                     "Ctx_cr" = "CTX CTRL", 
                                                     "Ctx_AxD" = "CTX AxD")
p4 <- DimPlot(seurat_filtered_markers, label = T, label.size = 4, group.by = "seurat_clusters", split.by = "ident") + theme(plot.title = element_blank()) + NoLegend()
ggsave("Supp_Fig6B_UMAPall.png", plot = p4, device = "png", width = 10, height = 4) 
```

Supp. Fig. 7A,B: RT-qPCR
```{r}
source("RTqPCR_plots.R")
```

# Proteomics
Correlation of RNAseq and proteomics data
t-test proteomic data were used for the correlation - diff. expr. proteins (p<0.05)
```{r}
# load proteomics data - differentially expressed proteins
setwd("..")
# diff. expressed proteins, q val(FDR) < 0.05
proteo <- read.csv("230619_proteomics/20230619_CVG_0516_MQ_proteome_complete_ttest_sign_gene_name_adjusted.txt", header = T, sep = "~")[,c(26,27,32)] #3459
colnames(proteo) <- c("qval","log2FC","gene")

# from those that are duplicated, choose the one with lower qvalue
proteo <- proteo %>% group_by(gene) %>% filter(qval == min(qval)) # 3423/3459 

# load seurat, prepare for DEA
seurat <- readRDS("cortical_integrated/seurat_cortical_integrated.rds")
seurat@active.assay <- "RNA"; seurat@active.ident <- as.factor(seurat$Condition)

# generate list of specific DEGs comparing cortical control and AxD, the gene list are differentially expressed proteins detected by proteomics, only those that can be found in both seurat and proteomics data (3356)
degs <- FindMarkers(seurat, ident.1 = "Ctx_AxD", ident.2 = "Ctx_cr", test.use = "t", assay = "RNA", 
                    logfc.threshold = 0, features = proteo[proteo$gene %in% rownames(seurat),]$gene, min.pct = 0) 
summary(degs)
degs <- degs %>% filter(is.na(p_val_adj) == FALSE & p_val_adj < 0.05) # 2521 genes
write.csv(degs, "230619_proteomics/DEGs_RNAseq.csv")

degs <- read.csv("230619_proteomics/DEGs_RNAseq.csv", header = T, row.names = 1)
# prepare data for the plot
proteo <- proteo %>% filter(gene %in% rownames(degs)) %>% arrange(gene) 
degs <- degs[order(rownames(degs)),]

df <- data.frame(gene = rownames(degs),
                 protein_diff = proteo$log2FC,
                 gene_log2FC = degs$avg_log2FC) 

gene.lab.myo <- df[grepl("^MYO",df$gene) | grepl("^MYH",df$gene) | grepl("^MYL",df$gene) | grepl("^MYT",df$gene),"gene"]
gene.lab.neuro <- c("NRXN1","DLX1","FOXG1","GRIA1","DCX","NRXN3","STMN2","MAP2","GRIA2","NCAN","NEFM","NEFL")
gene.others <- c("SOX2","GFAP","INA","NES","DES","LMNA","AQP4") #"LMNB1","LMNB2"
gene.proteasome <- read.table("230619_proteomics/proteasome_protein_list.txt", header = T)[,1]

# calculate correlation coefficient and statistics
coef <- cor(df$protein_diff, df$gene_log2FC) #Pearson 0.6698879
cor.stat <- cor.test(df$protein_diff, df$gene_log2FC, method = "pearson") #p-value < 2.2e-16
#cor(df$protein_diff, df$gene_log2FC, method = "spearman") #Spearman 0.67525
```

Fig. 6E: correlation plot
```{r}
# plot and save the two variables against each other with a line, add correlation info as annotation
p <- ggplot(df, aes(x=protein_diff, y=gene_log2FC, label=gene)) + 
    geom_hline(yintercept = 0, color = "#999999") +
    geom_vline(xintercept = 0, color = "#999999") +
    geom_point(color = "#cfcfcf") + 
    geom_point(data=subset(df, gene %in% gene.lab.myo), color = "#e06666", size = 3) +
    geom_point(data=subset(df, gene %in% gene.lab.neuro), color = "#639DD1", size = 3) +
    geom_point(data=subset(df, gene %in% gene.others), color = "#8FCE00", size = 3) +
    #geom_point(data=subset(df, gene %in% gene.proteasome), color = "#FFA500", size = 3) +
    geom_smooth(method = "lm", se = F, color = "black") +
    geom_text_repel(data=subset(df, gene %in% c(gene.lab.myo, gene.lab.neuro, gene.others)), max.overlaps = Inf) + 
  theme_bw() +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.margin = margin(r=0.5, b=0.5, l=0.5, t=0.5, unit = "cm")
  ) +
  xlab(expression("Protein log"[2]*"FC")) +
  ylab(expression("Gene log"[2]*"FC")) +
  annotate("text", x = -7, y = 5, size = 5, label = paste("r =", round(coef, digits = 3), "\n p-value < 2.2e-16"))
p

ggsave("230619_proteomics/Fig6E_correlation.png", plot = p, device = "png", width = 9, height = 6)
```

Intensities of selected proteins:
normalized data with imputed missing values we used in the intensity plots.
```{r}
setwd("..")
proteo <- read.delim("230619_proteomics/20231201_CVG_0516_D150_WholeProteome_Vulcano.txt", header = T, sep = "~")

# remove duplicates, leave the most significant 
proteo <- proteo %>% group_by(Gene.names) %>% filter(X.Log.P.value. == max(X.Log.P.value.)) # 6830/6926

goi <- list(GFAP = "GFAP", AQP4 = "AQP4", `SOX-2` = "SOX2", 
            `alpha-internexin` = "INA", `NEF-M` = "NEFM", `NEF-L` = "NEFL", FOXG1 = "FOXG1", nestin = "NES", desmin = "DES", `lamin-A/C` = "LMNA")

boxplot_protein <- function(gene, title){
  df <- proteo[proteo$Gene.names == gene,27:34] %>% t() %>% as.data.frame()
  colnames(df) <- "log2_intensity"
  df$log2_intensity <- as.numeric(df$log2_intensity)
  df$condition <- gsub(pattern = "LFQ.intensity.", rownames(df), replacement = "")
  df$condition <- gsub(pattern = "Ctrl", df$condition, replacement = "CTRL")
  df$condition <- substr(df$condition, start = 1, stop = nchar(as.character(df$condition)) - 2) %>% as.factor()
  
  p <- ggplot(df, aes(x = factor(df$condition, levels = c("CTRL","AxD")), y = log2_intensity, 
                  fill = factor(df$condition, levels = c("CTRL","AxD")))) + 
    geom_boxplot(linewidth = 1) + 
    geom_point(size = 3) +
    ylab(expression("Log"[2]* " imputed intensity")) +
    theme_light() +
    theme(axis.title.x = element_blank(),
          #axis.title.y = element_text(size = 14),
          axis.title.y = element_blank(),
          axis.text.x = element_text(size = 14, face = "bold", colour = "black"),
          axis.text.y = element_text(size = 14, colour = "black"),
          axis.line = element_line(linewidth = 1),
          axis.ticks = element_line(linewidth = 1, colour = "black"),
          legend.position = "none",
          plot.title = element_text(hjust = 0.5, face = "bold", size = 16)
        ) +
    scale_fill_manual(values = c("#639DD1","#E06666")) +
    ggtitle(title) +
    scale_y_continuous(limits = c(NA, 1.03*max(df$log2_intensity))) +
    stat_compare_means(method = "t.test", label = "p.signif", label.x = 1.5, size = 7, label.y = 1.005*max(df$log2_intensity), 
                       hide.ns = F, comparisons = list(c("CTRL","AxD")))

}

plots <- lapply(seq_along(goi), function(x) boxplot_protein(goi[[x]], names(goi)[x]))
```

Fig. 6F: Protein boxplots
```{r}
gridExtra::grid.arrange(grobs = plots, nrow=2)
ggsave("230619_proteomics/Fig6F_boxplot.png", plot = gridExtra::grid.arrange(grobs = plots, nrow=2), device = "png",
       width = 10, height = 7)
```

Supplementary Fig. 7E
```{r}
# volcano plot
proteo_volcano <- read.table("230619_proteomics/20231201_CVG_0516_D150_WholeProteome_Vulcano.txt", header = T, sep = "~")
proteo_volcano <- proteo_volcano %>% group_by(Gene.names) %>% filter(X.Log.P.value. == max(X.Log.P.value.))
proteo_volcano$pval <- 10^-(proteo_volcano$X.Log.P.value.)
res <- proteo_volcano[,c("pval","Difference","Gene.names")]
keyvals.colour <- ifelse(
  res$Difference < -0.65 & res$pval < 0.05, '#5A7BC6',
  ifelse(res$Difference > 0.65 & res$pval < 0.05, '#EB7321',
         '#E7E7E7'))
names(keyvals.colour)[keyvals.colour == '#5A7BC6'] <- 'downregulated'
names(keyvals.colour)[keyvals.colour == '#EB7321'] <- 'upregulated'
names(keyvals.colour)[keyvals.colour == '#E7E7E7'] <- 'insignificant'

gene.lab.neuro
gene.others
gene.proteasome
gene.lab.myo <- res[grepl("^MYO",res$Gene.names) | grepl("^MYH",res$Gene.names) | grepl("^MYL",res$Gene.names) | grepl("^MYT",res$Gene.names),]$Gene.names 
  
select.lab <-  res[abs(res$Difference) > 0.65 & res$pval < 0.05 & (res$Gene.names %in% c(gene.lab.neuro,gene.lab.myo,gene.others,gene.proteasome)),]$Gene.names 

p <- EnhancedVolcano(toptable = res,
                     lab = res$Gene.names,
                     selectLab = select.lab,
                     x = 'Difference',
                     y = 'pval',
                     axisLabSize = 8,
                     titleLabSize = 10,
                     title = "CTRL - AxD",
                     pCutoff = 0.05, 
                     colAlpha = 0.7,
                     colCustom = keyvals.colour,
                     subtitleLabSize = 9,
                     FCcutoff = 0.65, 
                     labSize = 3,
                     pointSize = 2.5,
                     drawConnectors = T,
                     maxoverlapsConnectors = Inf,
                     min.segment.length = 0.1,
                     arrowheads = FALSE,
                     boxedLabels = F,
                     cutoffLineType = "blank",
                     caption = "|log2FC| > 0.65 and p < 0.05") +
  theme_light() +
  theme(plot.subtitle = element_blank(), 
        plot.caption = element_text(size = 10), 
        plot.title = element_text(hjust = 0.5, size = 12),
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        axis.line = element_line(colour = "black", linewidth = 0.7, linetype = "solid"), 
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        axis.ticks = element_line(linewidth = 0.7, colour = "black"),
        axis.ticks.length = unit(0.2, "cm"),
        legend.position = "none") +
  xlab(expression("Log"[2]*"FC")) +
  guides(color = guide_legend(override.aes = list(size = 4))) +
  scale_y_continuous(breaks = c(0,5,10), limits = c(0,10))
p
ggsave(plot = p, "230619_proteomics/Supp_Fig7E_volcano.png", device = "png", width = 10, height = 7)
```

Supplementary Fig. 7F: GO enrichment on proteins
```{r}
#### ORA
background <- read.delim("230619_proteomics/20231201_CVG_0516_D150_WholeProteome_Vulcano.txt", header = T, sep = "~")[,"Gene.names"] %>% na.omit()

proteo <- read.csv("230619_proteomics/20230619_CVG_0516_MQ_proteome_complete_ttest_sign_gene_name_adjusted.txt", header = T, sep = "~")[,c(26,27,32)] #3459
colnames(proteo) <- c("qval","log2FC","gene")

DEA.list <- list("UP" = proteo[proteo$log2FC > 0.65 & proteo$qval < 0.05,],
                 "DOWN" = proteo[proteo$log2FC < -0.65 & proteo$qval < 0.05,])

ego.results <- list()

for (gene.list in 1:length(DEA.list)){    
  ego <- enrichGO(gene          = DEA.list[[gene.list]]$gene,
                  universe      = background,
                  keyType       = "ALIAS",
                  OrgDb         = org.Hs.eg.db,
                  ont           = "BP",
                  pAdjustMethod = "fdr",
                  pvalueCutoff  = 0.1,
                  qvalueCutoff  = 0.2,
                  readable      = TRUE,
                  minGSSize     = 5,
                  maxGSSize     = 800)
  ego <- simplify(ego)
  ego.results[[names(DEA.list)[gene.list]]] <- ego
}  
saveRDS(ego.results, "230619_proteomics/ORA_results.rds")

file.xlsx <- "230619_proteomics/ORA_results_sheet.xlsx"
write.xlsx("empty", file.xlsx, sheetName = "0", append = FALSE)
for (i in 1:length(ego.results)){
  if (dim(ego.results[[i]])[1] != 0) {
    write.xlsx(ego.results[[i]], file.xlsx, sheetName = names(ego.results)[i], append = TRUE)}
}
  
out <- merge_result(ego.results)
out@compareClusterResult <- out@compareClusterResult[order(out@compareClusterResult$p.adjust),]
out@compareClusterResult <- out@compareClusterResult[out@compareClusterResult$Count > 5,]
p <- dotplot(out, showCategory = 5, by = "GeneRatio", includeAll = T, label_format = 50) +
  theme_minimal() +
  theme(axis.text = element_text(size = 16, colour = "black"),
        axis.text.x = element_text(face = "bold"),
        axis.title.x = element_blank(),
        plot.title = element_text(size = 16, face = "bold", hjust = 1)) + 
  ggtitle("GO Terms Enrichment - Proteomics") +
  scale_color_gradient(low = "#8c0052",high = "#e0e0e0") 
p
ggsave(plot = p, filename = "230619_proteomics/Supp_Fig7F_ORA_plot.png", device = "png", bg = "white",
       width = 8, height = 5)
```

# Overlap with co-cultures
Supplemetary figure 8A-E
```{r}
source("overlap_with_coculture.R")
```
# Differential abundance
```{r}
library("scProportionTest") #scProportionTest_0.0.0.9000
proportion_test_func <- function(seurat.object, cluster_identity, sample_identity, plot_title){
  so <- readRDS(seurat.object)
  prop_test <- sc_utils(so)
  prop_test <- permutation_test(
	  prop_test, cluster_identity = cluster_identity,
	  sample_1 = levels(as.factor(so$Sample))[2], sample_2 = levels(as.factor(so$Sample))[1],
	  sample_identity = sample_identity, 
  )
  
  plot <- permutation_plot(prop_test) + theme(axis.text = element_text(size = 10),
                                              legend.position = "none") +
    ggtitle(plot_title)
  return(plot)
}

proportion_test_func(seurat.object = "cerebral/seurat_subset_cerebral.rds",
                     cluster_identity = "clusters_annot", sample_identity = "Sample", plot_title = "Cerebral organoids")
proportion_test_func(seurat.object = "cortical_integrated/seurat_cortical_integrated.rds",
                     cluster_identity = "clusters_annot", sample_identity = "Sample", plot_title = "Cortical organoids")
```
# Organoid quantification
Quantification plots of immunostaining in Fig. 5 and Fig. 6.
```{r org_quant}
dir <- dir("240610_figs_revision", pattern = "Fig6")
df.list2 <- list()
for (i in dir){
  df <- as.data.frame(read.table(paste0("240610_figs_revision/",i), header = T, sep = ""))
  name <- gsub(".txt","",i)
  name <- gsub("Fig6_","",name)
  colnames(df) <- c("sample","vals")
  df$sample <- factor(df$sample, levels = c("CTRL","AxD"))
  df.list2[[name]] <- df
} 

boxplot_custom <- function(df,x,y,fill,title){
  ggplot(data=df, aes(x=x,y=y,fill=fill)) + geom_boxplot(outlier.shape = NA) + geom_jitter(size = 2, width = 0.15) +
  theme_light() +
  theme(axis.title.x = element_blank(),
        legend.position = "none",
        axis.text = element_text(size = 16, colour = "black"),
        #axis.text.x = element_blank(),
        axis.line = element_line(linewidth = 0.5),
        axis.ticks = element_line(linewidth = 0.5, colour = "black"),
        axis.ticks.x = element_blank()
        ) +
  scale_fill_manual(values = c("#4472C4","#FFC000")) +
  ylab(title) +
  scale_y_continuous(expand = c(0,0), limits = c(0,1.2*max(y))) +
    # FOXG1 + SOX9 wilcox, DCX t-test
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x = 1.5, size = 5, label.y = 1.01*max(df[,2]), 
                      hide.ns = F, paired = F, comparisons = list(c("CTRL","AxD")))
}

plot.list2 <- lapply(seq_along(df.list2), function(x) boxplot_custom(df = df.list2[[x]], x = df.list2[[x]]$sample, y = df.list2[[x]]$vals, fill = df.list2[[x]]$sample, title = names(df.list2)[x]))

ggsave("Fig6_quantif_revision.svg", device = "svg", width = 5, height = 3, plot = cowplot::plot_grid(align = "v", ncol = 2, plotlist = plot.list2) %>% print())
```