---
title: "Co-cultures Figures"
author: "Zuzana"
date: '2022-10-06'
output: html_document
editor_options: 
  chunk_output_type: console
---
# Import libraries
```{r libraries}
library(Seurat)
library(Matrix)
library(SeuratWrappers)
#library(SoupX)
library(ggplot2)
library(scales)
library(DT)
library(dplyr)
library(dittoSeq)
#library(clustree)
library(VennDiagram)
library(data.table)
library(gridExtra)
library(RColorBrewer)
library(lemon)
library(EnhancedVolcano)
library(viridis)
library(ComplexHeatmap)
library(org.Hs.eg.db)
library(clusterProfiler)
library(xlsx)
library(enrichplot)
library(ggvenn)
library(reshape2)
```

```{r}
subfunction1 <- function(test){
  gene_list <- test$avg_log2FC * (-log10(test$p_val))
  names(gene_list) <- rownames(test)
  gene_list <- na.omit(gene_list)
  keep <- is.finite(gene_list)
  gene_list <- gene_list[keep]
  gene_list <- sort(gene_list, decreasing = TRUE)
  return(gene_list)
}
```

Set correct idents to iN/iA/precursors
```{r}
# neurons
seurat_iN <- readRDS("iN/seurat_iN_filtered_mt_rib.rds")
seurat_iN <- SetIdent(seurat_iN, value = seurat_iN$sct_pca_res.0.2)
seurat_iN <- RenameIdents(seurat_iN, "0" = "NEURO 1",
                                     "1" = "NEURO 2",
                                     "2" = "NEURO 3",
                                     "3" = "NEURO 4")
seurat_iN[["cell_type_3"]] <- seurat_iN@active.ident

# iA
seurat_iA <- readRDS("iPSC_iA/seurat_iA_filtered_mt_rib.rds")
seurat_iA <- SetIdent(seurat_iA, value = seurat_iA$cell_type_3)
seurat_iA <- RenameIdents(seurat_iA, "ASTRO_1" = "ASTRO 1",
                                     "ASTRO_2" = "ASTRO 2",
                                     "ASTRO_3" = "ASTRO 3",
                                     "ASTRO_4" = "ASTRO 4")
seurat_iA[["cell_type_4"]] <- seurat_iA@active.ident

# precursors
seurat_iPSC <- readRDS("iPSC_iA/seurat_iPSC_filtered_mt_rib.rds")

# set sample ident
seurat_celltype <- list("seurat_iN" = seurat_iN, "seurat_iA" = seurat_iA, "seurat_iPSC" = seurat_iPSC)
for (so in 1:length(seurat_celltype)){
  seurat_celltype[[so]] <- SetIdent(seurat_celltype[[so]], value = seurat_celltype[[so]]$Sample)
  # seurat_celltype[[so]] <- RenameIdents(seurat_celltype[[so]],
  #                    "Ctrl_wt" = "CTRL astroC/neuroC",
  #                    "Ctrl_w_a" = "CTRL astroAxD/neuroC",
  #                    "Ctrl_AxD" = "CTRL astroAxD/neuroAxD",
  #                    "OGD_wt" = "OGD astroC/neuroC",
  #                    "OGD_w_a" = "OGD astroAxD/neuroC",
  #                    "OGD_AxD" = "OGD astroAxD/neuroAxD"
  #                    )
    seurat_celltype[[so]] <- RenameIdents(seurat_celltype[[so]],
                     "Ctrl_wt" = "CTRL C/C",
                     "Ctrl_w_a" = "CTRL AxD/C",
                     "Ctrl_AxD" = "CTRL AxD/AxD",
                     "OGD_wt" = "OGD C/C",
                     "OGD_w_a" = "OGD AxD/C",
                     "OGD_AxD" = "OGD AxD/AxD"
                     )
  seurat_celltype[[so]][["Sample"]] <- seurat_celltype[[so]]@active.ident
}

seurat_iN <- seurat_celltype[["seurat_iN"]]
seurat_iN <- SetIdent(seurat_iN, value = seurat_iN$cell_type_3)
seurat_iA <- seurat_celltype[["seurat_iA"]]
seurat_iA <- SetIdent(seurat_iA, value = seurat_iA$cell_type_4)
seurat_iPSC <- seurat_celltype[["seurat_iPSC"]]
seurat_iPSC <- SetIdent(seurat_iPSC, value = seurat_iPSC$cell_type_5)
```


# Fig 1
1B: UMAP
```{r}
seurat.object <- readRDS("seurat_all_merged_filtered_mt_rib.rds")
DimPlot(seurat.object, group.by = "cell_type_joined", pt.size = 1)
seurat.object@active.ident <- seurat.object$cell_type_joined
seurat.object <- RenameIdents(seurat.object, 
                              "EPCAM/CDH1" = "AxD cluster",
                              "PRE-ASTRO" = "AxD cluster",
                              "PRE-NEURO" = "AxD cluster",
                              "DCX/STMN1" = "AxD cluster",
                              "ASTRO_1" = "ASTRO 1",
                              "ASTRO_2" = "ASTRO 2",
                              "ASTRO_3" = "ASTRO 3",
                              "ASTRO_4" = "ASTRO 4",
                              "NEURO_1" = "NEURO 1",
                              "NEURO_2" = "NEURO 2",
                              "NEURO_3" = "NEURO 3",
                              "NEURO_4" = "NEURO 4"
                              )
myCol <- c("#008000","#C6DBEF", "#6BAED6", "#2171B5", "#08306B","#FCBBA1", "#FC9272", "#EF3B2C", "#A50F15")

seurat.object$clusters_1B <- Idents(seurat.object)

p <- DimPlot(seurat.object, pt.size = 0.8, cols = myCol, label = F) + 
  theme(legend.position = "none")
ggsave("Fig1B_UMAPall.png", plot = p, device = "png", height = 6, width = 7)

table(seurat.object@active.ident)
```

1C: Dotplot of markers
```{r}
seurat.object <- readRDS("seurat_all_merged_filtered_mt_rib.rds")
seurat.object <- RenameIdents(seurat.object,
                              "ASTRO 1" = "Astrocytes",
                              "ASTRO 2" = "Astrocytes",
                              "ASTRO 3" = "Astrocytes",
                              "ASTRO 4" = "Astrocytes",
                              "NEURO 1" = "Neurons", 
                              "NEURO 2" = "Neurons", 
                              "NEURO 3" = "Neurons", 
                              "NEURO 4" = "Neurons")
  
seurat.object$clusters_1C <- seurat.object@active.ident
seurat.object@active.assay <- "RNA"

#saveRDS(seurat.object, "seurat_all_merged_filtered_mt_rib.rds")

# markers <- FindAllMarkers(seurat.object, min.pct = 0.1, logfc.threshold = 0.25, only.pos = TRUE)
# markers <- markers[markers$avg_log2FC > 1 & markers$p_val_adj < 0.05,]
# write.csv(markers, "Fig1_broad_markers.csv")

neuro <- c("GAP43","NEFL","NEFM","MAP2","DCX")
astro <- c("VIM","S100B","FABP7","GFAP","CRYAB")
precursors <- c("EPCAM","CDH1","CLDN7","KRT18","KRT8")
gene.set <- c(precursors, astro, neuro)
#gene.set <- unique(gene.set)

seurat.object@active.ident <- factor(seurat.object@active.ident, levels = c("AxD cluster","Astrocytes","Neurons"))

p.dot <- DotPlot(seurat.object, features = gene.set) +
  scale_color_gradient(high = "#8c0052", low = "#e0e0e0") +
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle=60, vjust = 1, hjust =1),
        legend.title = element_text(size = 10), legend.text = element_text(size = 10), 
        legend.key.size = unit(0.5, "cm"), legend.text.align = 1, legend.title.align = 0, 
        plot.margin = margin(r = 1, l = 0.25, t = 0.25, b = 0.25, unit = "cm"),
        legend.margin = margin(t = 0, r = 0, l = 0, b = -0.25, unit = "cm"),
        plot.title = element_text(size = 18,face = "plain", hjust = 0.5)) +
  guides(colour = guide_colorbar(order = 1, title = "Scaled Average\nExpression"), 
         size = guide_legend(order = 2, title = "Percent\nExpressed")) +
  coord_flip()
p.dot
ggsave(p.dot,filename = "Fig1C_markers.png",device = "png", bg="white", width = 3.5, height = 5)
```

1D: Proportional barplot, main populations
```{r}
seurat.object <- SetIdent(seurat.object, value = seurat.object$Sample)
seurat.object <- RenameIdents(seurat.object,
                              "Ctrl_wt" = "CTRL astroC/neuroC",
                              "Ctrl_w_a" = "CTRL astroAxD/neuroC",
                              "Ctrl_AxD" = "CTRL astroAxD/neuroAxD",
                              "OGD_wt" = "OGD astroC/neuroC",
                              "OGD_w_a" = "OGD astroAxD/neuroC",
                              "OGD_AxD" = "OGD astroAxD/neuroAxD")
seurat.object$Sample <- seurat.object@active.ident

myCol2 <- c("#00A7F3","#EC1C24","#0D9433")

p <- dittoBarPlot(seurat.object, var = seurat.object$clusters_1C, group.by = "Sample", data.out = T)
p[["data"]]$grouping <- factor(p[["data"]]$grouping, levels = c("OGD astroAxD/neuroAxD","OGD astroAxD/neuroC","OGD astroC/neuroC","CTRL astroAxD/neuroAxD","CTRL astroAxD/neuroC","CTRL astroC/neuroC"))
p[["data"]]$percent100 <- p[["data"]]$percent * 100
data <- p[["data"]]

p.bar <- ggplot(data = data, aes(x = percent100, y = grouping, fill = label)) + geom_bar(stat = "identity",width = 0.8) +
  theme_void() +
  theme(axis.text.x = element_text(size = 12, face = "bold"),
        axis.line.y = element_blank(), axis.ticks.y = element_blank(),
        axis.title.x = element_text(size=12, vjust = -1.5),
        axis.line.x = element_line(linewidth = 1, lineend = "square"),
        axis.text.y = element_text(size=12, face = "bold", margin = margin(r = 0.2, unit = "cm"), hjust = 1), 
        axis.ticks.x = element_line(size = 1),
        axis.ticks.length.x = unit(0.25, "cm"),
        plot.margin = margin(0,0,10,2, unit = "pt"),
        legend.position = "none",
        plot.title = element_blank()) +
  coord_capped_cart(bottom = "both") +
  scale_fill_manual(values = myCol2) +
  xlab("Percent of Cells") +
  scale_x_continuous(breaks = c(0,50,100)) 
p.bar
ggsave("Fig1D_barplotall.png", device = "png",bg="white", plot = p.bar, width = 5, height = 2)

table(seurat.object$Sample)
```

1E: Proportional barplot, subpopulations
```{r}
# astrocytes
myCol.a <- c("#C6DBEF", "#6BAED6", "#2171B5", "#08306B")
p.a <- dittoBarPlot(seurat_iA, var = seurat_iA@active.ident, group.by = "Sample", data.out = T)
p.a[["data"]]$grouping <- factor(p.a[["data"]]$grouping, levels = c("OGD astroAxD/neuroAxD","OGD astroAxD/neuroC","OGD astroC/neuroC","CTRL astroAxD/neuroAxD","CTRL astroAxD/neuroC","CTRL astroC/neuroC"))
p.a[["data"]]$percent100 <- p.a[["data"]]$percent * 100
data.a <- p.a[["data"]]

p.a.bar <- ggplot(data = data.a, aes(x = percent100, y = grouping, fill = label)) + geom_bar(stat = "identity",width = 0.8) +
  theme_void() +
  theme(axis.text.x = element_text(size = 12, face = "bold"),
        axis.line.y = element_blank(), axis.ticks.y = element_blank(),
        axis.title.x = element_text(size=12, vjust = -1.5, margin = margin(0,0,10,0, unit = "pt")),
        axis.line.x = element_line(size = 1, lineend = "square"),
        axis.text.y = element_text(size=12, face = "bold", margin = margin(r = 0.2, unit = "cm"), hjust = 1), 
        axis.ticks.x = element_line(size = 1),
        axis.ticks.length.x = unit(0.25, "cm"),
        plot.margin = margin(l=2,r=0,b=10,t=10, unit = "pt"), 
        legend.position = "bottom", legend.direction = "horizontal",
        legend.key.size = unit(0.3,"cm"), legend.title = element_blank(),
        legend.text = element_text(size = 10),
        plot.title = element_text(size = 14, hjust = 0.5, vjust = 4)) +
  coord_capped_cart(bottom = "both") +
  scale_fill_manual(values = myCol.a) +
  xlab("Percent of Cells") +
  scale_x_continuous(breaks = c(0,50,100)) +
  ggtitle("iA clusters")
p.a.bar
#ggsave("Fig1E_barplot_iA.png", device = "png",bg="white", plot = p.a.bar, width = 5, height = 3)

# neurons
seurat_iN@active.ident <- seurat_iN$cell_type_3
myCol.n <- c("#FCBBA1", "#FC9272", "#EF3B2C", "#A50F15")
p.n <- dittoBarPlot(seurat_iN, var = seurat_iN@active.ident, group.by = "Sample", data.out = T)
p.n[["data"]]$grouping <- factor(p.n[["data"]]$grouping, levels = c("OGD astroAxD/neuroAxD","OGD astroAxD/neuroC","OGD astroC/neuroC","CTRL astroAxD/neuroAxD","CTRL astroAxD/neuroC","CTRL astroC/neuroC"))
p.n[["data"]]$percent100 <- p.n[["data"]]$percent * 100
data.n <- p.n[["data"]]

p.n.bar <- ggplot(data = data.n, aes(x = percent100, y = grouping, fill = label)) + geom_bar(stat = "identity",width = 0.8) +
  theme_void() +
  theme(axis.text.x = element_text(size = 12, face = "bold"),
        axis.line.y = element_blank(), axis.ticks.y = element_blank(),
        axis.title.x = element_text(size=12, vjust = -1.5, margin = margin(0,0,10,0, unit = "pt")),
        axis.line.x = element_line(size = 1, lineend = "square"),
        axis.text.y = element_text(size=12, face = "bold", margin = margin(r = 0.2, unit = "cm")), 
        axis.ticks.x = element_line(size = 1),
        axis.ticks.length.x = unit(0.25, "cm"),
        plot.margin = margin(l=2,r=0,b=10,t=10, unit = "pt"), 
        legend.position = "bottom", legend.direction = "horizontal", 
        legend.key.size = unit(0.3,"cm"), legend.title = element_blank(),
        legend.text = element_text(size = 10),
        plot.title = element_text(size = 14, hjust = 0.5, vjust = 4)) +
  coord_capped_cart(bottom = "both") +
  scale_fill_manual(values = myCol.n) +
  xlab("Percent of Cells") +
  scale_x_continuous(breaks = c(0,50,100)) +
  ggtitle("iN clusters")
p.n.bar
#ggsave("Fig1E_barplot_iN.png", device = "png",bg="white", plot = p.n.bar, width = 5, height = 3)

ggsave("Fig1E_barplot_iA_iN.png", plot = egg::ggarrange(p.a.bar, p.n.bar + theme(axis.text.y = element_blank()), nrow = 1), device = "png", height = 3, width = 9.5)
```

1F: Violin GFAP
```{r}
seurat_iA <- readRDS("iPSC_iA/seurat_iA_filtered_mt_rib.rds")
myCol.a <- c("#C6DBEF", "#6BAED6", "#2171B5", "#08306B")
p <- VlnPlot(seurat_iA, features = "GFAP", cols = myCol.a, pt.size = 0) + NoLegend() +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 0, size = 16),
        axis.title.y = element_blank(),
        axis.text.y = element_text(size = 16)) +
  scale_x_discrete(limits = rev)
ggsave("Fig1F_violin.png", plot = p, device = "png", height = 4, width = 3.5)
```

1G: ORA AxD cells
```{r}
seurat.object@active.ident <- seurat.object$clusters_1B
seurat.object@active.assay <- "RNA"
markers.precursors <- FindMarkers(seurat.object, ident.1 = "AxD cluster", min.pct = 0.1, logfc.threshold = 0.25)
markers.precursors <- markers.precursors[abs(markers.precursors$avg_log2FC) > 1 & markers.precursors$p_val_adj < 0.05,]
write(rownames(markers.precursors[markers.precursors$avg_log2FC > 1,]), "markers_precursors_UP.txt")
write(rownames(markers.precursors[markers.precursors$avg_log2FC < -1,]), "markers_precursors_DOWN.txt")

reg <- c("UP","DOWN")
background <- rownames(seurat.object)
ego.all <- list()
for(i in reg){
  gene.list <- as.vector(read.table(paste0("markers_precursors_",i,".txt"), header = F)[,1]) 
      
  ego <- enrichGO(gene       = gene.list,
               universe      = background,
               keyType       = "ALIAS",
               OrgDb         = org.Hs.eg.db,
               ont           = "ALL",
               pAdjustMethod = "fdr",
               pvalueCutoff  = 0.1,
               qvalueCutoff  = 0.2,
               readable      = TRUE,
               minGSSize     = 3)
  
  ego <- simplify(ego)
  ego.all[[i]] <- ego
}
saveRDS(ego.all, "precursors_ORA.rds")

ego.all <- readRDS("Fig1_precursors_ORA.rds")
out <- merge_result(ego.all)
out@compareClusterResult <- out@compareClusterResult[order(out@compareClusterResult$p.adjust),]
out@compareClusterResult <- out@compareClusterResult[out@compareClusterResult$Count > 5,]
p <- dotplot(out, showCategory = 5, by = "GeneRatio", includeAll = F, label_format = 50) +
  scale_color_gradient(low = "#8c0052", high = "#e0e0e0") +
  theme_minimal() +
  theme(axis.text = element_text(size = 16, colour = "black"),
        axis.text.x = element_text(face = "bold"),
        axis.title.x = element_blank(),
        plot.title = element_text(size = 16, face = "bold", hjust = 0.75)) + 
  ggtitle("GO Terms Enrichment in Precursors")
p
ggsave(plot = p, filename = "Fig1G_AxDcluster_ORA_dotplot.png", device = "png", bg = "white", height = 4.5, width = 8)
```

# Fig 2B,D,E
```{r}
df.list <- list()
for (i in dir){
  df <- as.data.frame(read.table(paste0("240610_figs_revision/",i), header = T, sep = ""))
  name <- substr(i,7,nchar(i)-15)
  colnames(df) <- c("sample","experiment","vals")
  df$sample <- factor(df$sample, levels = c("wt","wa"))
  df.list[[name]] <- df
} 

boxplot_custom <- function(df,x,y,fill,title){
  ggplot(data=df, aes(x=x,y=y,fill=fill)) + geom_boxplot(outlier.shape = NA) + geom_jitter(size = 1, width = 0.15) +
  theme_light() +
  theme(axis.title.x = element_blank(),
        legend.position = "none",
        axis.text = element_text(size = 16, colour = "black"),
        axis.text.x = element_blank(),
        axis.line = element_line(linewidth = 0.5),
        axis.ticks = element_line(linewidth = 0.5, colour = "black"),
        axis.ticks.x = element_blank()
        ) +
  scale_fill_manual(values = c("#4472C4","#FFC000")) +
  ylab(title) +
  scale_y_continuous(expand = c(0,0), limits = c(0,1.2*max(y))) +
  stat_compare_means(method = "t.test", label = "p.signif", label.x = 1.5, size = 5, label.y = 1.01*max(df[,3]), 
                      hide.ns = F, paired = F, comparisons = list(c("wt","wa")))
}

plot.list1 <- lapply(seq_along(df.list), function(x) boxplot_custom(df = df.list[[x]], x = df.list[[x]]$sample, y = df.list[[x]]$vals, fill = df.list[[x]]$sample, title = names(df.list)[x]))
grid.arrange(grobs = plot.list1)

# 2E
df.list2 <- list()
for (i in dir){
  df <- as.data.frame(read.table(paste0("240610_figs_revision/",i), header = T, sep = ""))
  name <- substr(i,7,nchar(i)-9)
  colnames(df) <- c("sample","vals")
  df$sample <- factor(df$sample, levels = c("wt","wa"))
  df.list2[[name]] <- df
} 

boxplot_custom <- function(df,x,y,fill,title){
  ggplot(data=df, aes(x=x,y=y,fill=fill)) + geom_boxplot(outlier.shape = NA) + geom_jitter(size = 1, width = 0.15) +
  theme_light() +
  theme(axis.title.x = element_blank(),
        legend.position = "none",
        axis.text = element_text(size = 16, colour = "black"),
        axis.text.x = element_blank(),
        axis.line = element_line(linewidth = 0.5),
        axis.ticks = element_line(linewidth = 0.5, colour = "black"),
        axis.ticks.x = element_blank()
        ) +
  scale_fill_manual(values = c("#4472C4","#FFC000")) +
  ylab(title) +
  scale_y_continuous(expand = c(0,0), limits = c(0,1.2*max(y))) +
  stat_compare_means(method = "wilcox.test", label = "p.format", label.x = 1.5, size = 5, label.y = 1.01*max(df[,2]), 
                      hide.ns = F, paired = F, comparisons = list(c("wt","wa")))
}

plot.list2 <- lapply(seq_along(df.list2), function(x) boxplot_custom(df = df.list2[[x]], x = df.list2[[x]]$sample, y = df.list2[[x]]$vals, fill = df.list2[[x]]$sample, title = names(df.list2)[x]))

plot.list <- c(plot.list1,plot.list2)
ggsave("Fig2_barplots_revision.svg", device = "svg", width = 6, height = 5, plot = cowplot::plot_grid(align = "v", nrow = 2, plotlist = plot.list) %>% print())
```

# Fig 3
3A: Volcano iA and iN: astroC/neuroC - astroAxD/neuroC
```{r}
DE_tests <- readRDS("iPSC_iA/DEA_mt_rib/DE_iA.rds")
DE_tests <- readRDS("iN/DEA_mt_rib/DE_iN.rds")
res <- DE_tests[["Ctrl_wt-Ctrl_w_a"]]

keyvals.colour <- ifelse(
  res$avg_log2FC < -0.65 & res$p_val_adj < 0.05, '#5A7BC6',
  ifelse(res$avg_log2FC > 0.65 & res$p_val_adj < 0.05, '#EB7321',
         '#E7E7E7')) 
names(keyvals.colour)[keyvals.colour == '#5A7BC6'] <- 'downregulated'
names(keyvals.colour)[keyvals.colour == '#EB7321'] <- 'upregulated'
names(keyvals.colour)[keyvals.colour == '#E7E7E7'] <- 'insignificant'

top10up <- res[res$avg_log2FC > 0.65 & res$p_val_adj < 0.05,]
top10up <- rownames(head(top10up[order(top10up$p_val_adj, decreasing = F),],10))
top10down <- res[res$avg_log2FC < -0.65 & res$p_val_adj < 0.05,]
top10down <- rownames(head(top10down[order(top10down$p_val_adj, decreasing = F),],10))

p <- EnhancedVolcano(toptable = res,
                     lab = rownames(res),
                     selectLab = c(top10up,top10down,"GFAP"),
                     x = 'avg_log2FC',
                     y = 'p_val_adj',
                     title = "iA: astroC/neuroC - astroAxD/neuroC",
                     pCutoff = 0.05, 
                     colAlpha = 0.7,
                     colCustom = keyvals.colour,
                     FCcutoff = 0.65, 
                     labSize = 4.5,
                     pointSize = 3,
                     drawConnectors = T,
                     maxoverlapsConnectors = Inf,
                     min.segment.length = 0.1,
                     cutoffLineType = "blank",
                     arrowheads = F,
                     caption = "|log2FC| > 0.65 and p-adj < 0.05") +
  theme_light() +
  theme(plot.subtitle = element_blank(), 
        plot.caption = element_text(size = 10), 
        plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        axis.line = element_line(colour = "black", linewidth = 0.7, linetype = "solid"), 
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        axis.ticks = element_line(linewidth = 0.7, colour = "black"),
        axis.ticks.length = unit(0.2, "cm"),
        legend.position = "none") +
  xlab(expression("Log"[2]*"FC")) +
  guides(color = guide_legend(override.aes = list(size = 4)))
p
ggsave(plot = p, "Fig3A_volcano_iA.png", device = "png", width = 6, height = 5)
ggsave(plot = p, "Fig3A_volcano_iN.png", device = "png", width = 6, height = 5)
```

3B: Heatmap 
```{r}
DE_tests <- list("iA" = readRDS("iPSC_iA/DEA_mt_rib/DE_iA.rds"),
                 "iN" = readRDS("iN/DEA_mt_rib/DE_iN.rds"))
goi_list <- list()
for (i in 1:length(DE_tests)){
  gene.lists.up <- lapply(DE_tests[[i]][c("Ctrl_wt-Ctrl_AxD","Ctrl_wt-Ctrl_w_a")], 
                    function(x) x = x[(x$avg_log2FC > 0.65) & x$p_val_adj < 0.05,])
  gene.lists.down <- lapply(DE_tests[[i]][c("Ctrl_wt-Ctrl_AxD","Ctrl_wt-Ctrl_w_a")], 
                     function(x) x = x[(x$avg_log2FC < -0.65) & x$p_val_adj < 0.05,])
  reference.gene.set.up <- Reduce(intersect,lapply(gene.lists.up, function(x) rownames(x)))
  reference.gene.set.down <- Reduce(intersect,lapply(gene.lists.down, function(x) rownames(x)))
  reference.gene.set <- c(reference.gene.set.up,reference.gene.set.down)
  
  goi_list[[names(DE_tests)[i]]] <- reference.gene.set
}

plot_data <- list() 
seurat_celltype <- list("iA" = readRDS("iPSC_iA/seurat_iA_filtered_mt_rib.rds"),
                        "iN" = readRDS("iN/seurat_iN_filtered_mt_rib.rds"))

for (i in 1:(length(seurat_celltype))){
  seurat_object <- seurat_celltype[[i]]
  DefaultAssay(seurat_object) <- "RNA"
  seurat_object <- SetIdent(seurat_object, value = "Sample")
  seurat_object <- subset(seurat_object, ident = c("Ctrl_wt", "Ctrl_w_a", "Ctrl_AxD"))
  seurat_object <- RenameIdents(seurat_object, "Ctrl_wt" = "astroC/neuroC",
                                     "Ctrl_w_a" = "astroAxD/neuroC", 
                                     "Ctrl_AxD" = "astroAxD/neuroAxD")
 
  all.genes <- rownames(seurat_object)
  seurat_object <- NormalizeData(seurat_object, normalization.method = "LogNormalize", scale.factor = 10000, assay = "RNA")
  seurat_object <- ScaleData(seurat_object, features = all.genes, assay = "RNA")
  
  averages <- AverageExpression(seurat_object, return.seurat=TRUE)
  plot_data[[i]] <- averages
}

plot_list <- lapply(seq_along(plot_data), function(x) DoHeatmap(plot_data[[x]], features = goi_list[[x]], draw.lines = F, group.bar = T,
                                        angle = 0, group.bar.height = 0, hjust = 0.5, size = 0, 
                                        group.colors = c("white","white","white")) + 
         scale_fill_gradientn(colors = c("#E7E7E7","#EB7321")) + 
         theme(axis.text.y = element_text(size = 15, color = "black"),
               axis.text.x = element_text(size = 12, color = "black", angle = 90, vjust = 0.5, hjust = 1)) +
         coord_flip() +
         scale_x_discrete(limits = rev) +
         guides(colour = "none")
       )

egg::ggarrange(plot_list[[1]],plot_list[[2]], nrow = 1, common.legend = T)

ggsave("Fig3C_heatmap_iA.png", plot = p, device = "png", width = 14, height = 3)
ggsave("Fig3_heatmap_iN.png", plot = p, device = "png", width = 7, height = 3)
```

3C: Interaction number and interaction strength difference
```{r}
cir1 <- netVisual_diffInteraction(cellchat.merged, weight.scale = T, arrow.size = 0.5, margin = 0.1)
cir2 <- netVisual_diffInteraction(cellchat.merged, weight.scale = T, arrow.size = 0.5, margin = 0.1,
                                  measure = "weight")
svg("Fig3D_CellChat_circle.svg")
  par(mfrow= c(2,1))
  cir1
  cir2
dev.off()
```

3D: Information flow chart 
```{r}
gg6 <- rankNet(cellchat.merged, mode = "comparison", stacked = T, do.stat = TRUE, color.use = c("#7AA3CD","#D17480"))
ggsave("Fig3E_CellChat_infflow.png",plot = gg6 + theme(legend.position = "none"), device = "png", width = 3, height = 5.5)
```

3E: Outgoing and incoming signaling in selected pathways
```{r}
object.list <- list("Ctrl_wt" = cellchat1, "Ctrl_w_a" = cellchat2)
pathways.show <- c("FGF","TGFb","LIFR","ncWNT","NOTCH","EGF","RELN")
i <- 1
ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "outgoing", signaling = pathways.show, title = names(object.list)[i], width = 6, height = 6)
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "outgoing", signaling = pathways.show, title = names(object.list)[i+1], width = 6, height = 6)

svg("Fig3F_CellChat_outgoing.svg")
ComplexHeatmap::draw(ht1 + ht2, ht_gap = unit(1, "cm"))
dev.off()

ht3 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "incoming", signaling = pathways.show, title = names(object.list)[i], width = 6, height = 6, color.heatmap = "GnBu")
ht4 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "incoming", signaling = pathways.show, title = names(object.list)[i+1], width = 6, height = 6, color.heatmap = "GnBu")

svg("Fig3F_CellChat_incoming.svg")
ComplexHeatmap::draw(ht3 + ht4, ht_gap = unit(1, "cm"))
dev.off()
```

# Fig 4
4B: Barplot
```{r}
df3 <- read.table("Fig4B_ldh_revised.txt", header = T, sep = "")
df3$sample <- factor(df3$sample, levels = c("wt","wa"))
df3$treatment <- factor(df3$treatment, levels = c("ctrl","ogd"))

p <- ggplot(data=df3, aes(x=sample,y=vals,fill=treatment)) + geom_boxplot(outlier.shape = NA, position = position_dodge(preserve = "single")) + geom_point(position=position_jitterdodge(dodge.width=0.7, jitter.width = 0.15)) +
  theme_light() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_text(size = 16),
        legend.position = c(0.15,0.85),
        legend.text = element_text(size = 14),
        legend.title = element_blank(),
        axis.text = element_text(size = 16, colour = "black"),
        axis.line = element_line(linewidth = 1),
        axis.ticks = element_line(linewidth = 1, colour = "black")
        ) +
  scale_fill_manual(values = c("#4472C4","#FF0000")) +
  scale_y_continuous(expand = c(0,0), limits = c(0,1.2*max(df3$vals))) +
  ylab("Cell death (% LDH release)")
p
ggsave("Fig4B_barplots.svg", plot = p, device = "svg", width = 4, height = 3)

```

4C: Table Heatmap
```{r}
DE_iA <- readRDS("iPSC_iA/DEA_mt_rib/DE_iA.rds")
DE_iN <- readRDS("iN/DEA_mt_rib/DE_iN.rds")
DE_iPSC <- readRDS("iPSC_iA/DEA_mt_rib/DE_iPSC.rds")
DE.list <- list(DE_iA,DE_iN,DE_iPSC)

DEGs <- lapply(DE.list, function(x) x = lapply(x[c("Ctrl_wt-OGD_wt","Ctrl_w_a-OGD_w_a")], function(y) y = y[(abs(y$avg_log2FC) > 0.65) & y$p_val_adj < 0.05,]))
names(DEGs) <- c("iA","iN","precursors")

DEG.num <- unlist(lapply(DEGs, function(x) x = sapply(x, FUN = nrow)))

dt <- cbind(c(0,4),c(0,8),c(1,14))
colnames(dt) <- c("iN","iA","AxD cluster")
rownames(dt) <- c("astroC/neuroC","astroAxD/neuroC")

dt <- melt(dt)
dt$rel_value = dt$value/max(dt$value)

htmap <- dt %>%
    group_by(Var2) %>%
    ggplot(aes(Var2,Var1)) +
    geom_tile(aes(fill = rel_value), colour = "white") +
    geom_text(aes(label = value), size = 8) +
    scale_fill_gradient2(low = "#FCECEA", mid = "#f1a099", high = "#e34234", midpoint = 0.7) +
    theme(panel.grid.major = element_blank(),
                panel.grid.minor = element_blank(),
                panel.background = element_rect(fill = "white"),
                plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
                axis.text.y = element_text(size = 14, face = "bold", colour = "black"),
                axis.text.x = element_text(hjust = 1, vjust = 1, size = 14, 
                                           face = "bold", colour = "black", angle = 30),
                axis.title = element_blank(),
                axis.ticks = element_blank()) +
    ggtitle("Number of DEGs") +
    labs(fill = "Heatmap") +
  NoLegend() +
  scale_x_discrete(position = "bottom") +
  coord_flip()
htmap
ggsave("Fig4C_heatmap.png",plot=htmap, device = "png", width = 3, height = 4)
```

4D: Information flow chart - corrected CTRL x OGD
```{r}
cellchat1 <- readRDS("CellChat/CTRL/Ctrl_wt_cellchat_processed.rds")
cellchat2 <- readRDS("CellChat/OGD/OGD_wt_cellchat_processed.rds")
cellchat.merged <- mergeCellChat(list(cellchat1,cellchat2), add.names = c("Ctrl_wt","OGD_wt"))

gg6 <- rankNet(cellchat.merged, mode = "comparison", stacked = T, do.stat = TRUE, color.use = c("#7AA3CD","#D17480"))
ggsave("Fig4D_CellChat_infflow.png",plot = gg6 + theme(legend.position = "none"), device = "png", width = 3, height = 5.5)
```

4E: Information flow chart - astroAxD/neuroC CTRL x OGD
```{r}
cellchat1 <- readRDS("CellChat/OGD/CTRL_w_a_processed_woAS4.rds")
cellchat2 <- readRDS("CellChat/OGD/OGD_w_a_cellchat_processed.rds")
cellchat.merged <- mergeCellChat(list(cellchat1,cellchat2), add.names = c("Ctrl_w_a","OGD_w_a"))

gg6 <- rankNet(cellchat.merged, mode = "comparison", stacked = T, do.stat = TRUE, color.use = c("#7AA3CD","#D17480"))
ggsave("Fig4E_CellChat_infflow.png",plot = gg6 + theme(legend.position = "none"), device = "png", width = 3, height = 5.5)
```

4F: Outgoing and incoming signaling in selected pathways, astroAxD/neuroC OGD
```{r}
cellchat2 <- readRDS("CellChat/OGD/OGD_w_a_cellchat_processed.rds")
pathways.show <- c("ACTIVIN","TGFb","BMP","NRG")

ht1 = netAnalysis_signalingRole_heatmap(cellchat2, pattern = "outgoing", signaling = pathways.show, title = "astroAxD/neuroC OGD - outgoing", width = 6, height = 4)

svg("Fig4F_CellChat_outgoing.svg")
ComplexHeatmap::draw(ht1)
dev.off()

ht3 = netAnalysis_signalingRole_heatmap(cellchat2, pattern = "incoming", signaling = pathways.show, title = "astroAxD/neuroC OGD - incoming", width = 6, height = 4, color.heatmap = "GnBu")

svg("Fig4F_CellChat_incoming.svg")
ComplexHeatmap::draw(ht3)
dev.off()
```

4G,H: PDGF signaling in OGD astroC/neuroC
```{r}
cellchat2 <- readRDS("CellChat/OGD/OGD_wt_cellchat_processed.rds")

pathways.show <- c("PDGF","NGL","NEGR")
svg("Fig4G_CellChat_wt.svg")
  netVisual_chord_gene(cellchat2, signaling = pathways.show, sources.use = c(1:9), 
                       targets.use = c(1:9), lab.cex = 1.5,legend.pos.y = 15, 
                       legend.pos.x = 10,big.gap = 20, small.gap = 10)
dev.off()

svg("Fig4H_CellChat_PDGF.svg")
  netAnalysis_signalingRole_network(cellchat2, signaling = "PDGF", width = 8, height = 4, font.size = 10)
dev.off()
```

#Supplement 1
Supp 1B
```{r}
seurat_integrated <- readRDS("seurat_integrated.rds")
seurat_integrated <- SetIdent(seurat_integrated, value = seurat_integrated$Sample)
seurat_integrated <- RenameIdents(seurat_integrated,
                   "Ctrl_wt" = "CTRL astroC/neuroC",
                   "Ctrl_w_a" = "CTRL astroAxD/neuroC",
                   "Ctrl_AxD" = "CTRL astroAxD/neuroAxD",
                   "OGD_wt" = "OGD astroC/neuroC",
                   "OGD_w_a" = "OGD astroAxD/neuroC",
                   "OGD_AxD" = "OGD astroAxD/neuroAxD"
                   )
seurat_integrated[["Sample"]] <- seurat_integrated@active.ident

vlnplotCustom <- function(features, intercept){
  return(VlnPlot(seurat_integrated, features = features, pt.size = 0, group.by = "Sample") +
           geom_hline(yintercept = intercept) +
           NoLegend() +
           theme(axis.title.x = element_blank(),
                   plot.title = element_text(size = 12)))
}

ggsave("Supp_Fig1B_VlnQC.png", plot = egg::ggarrange(nrow = 3, #cowplot::plot_grid(align = "hv", nrow = 3, axis = "b", 
  vlnplotCustom(features = "nCount_RNA", intercept = 2500) + theme(axis.text.x = element_blank()),
  vlnplotCustom(features = "nFeature_RNA", intercept = 1500) + theme(axis.text.x = element_blank()),
  vlnplotCustom(features = "percent.mt", intercept = 10) + theme(axis.text.x = element_text(size = 10))
), device = "png", width = 5, height = 7)
```

Supp 1C
```{r}
seurat_annotated <- readRDS("seurat_annotated.rds")
p4 <- DimPlot(seurat_annotated, group.by = "sct_pca_res.0.4", label = T, label.size = 6) + theme(plot.title = element_blank()) + NoLegend()
ggsave("Supp_Fig1C_UMAP.png", plot = p4, device = "png", width = 5, height = 5) 
```

Supp 1D
```{r}
p5 <- FeaturePlot(seurat_annotated, features = c("percent.mt","MALAT1"), min.cutoff = "q70", max.cutoff = "q90",cols = c("#e0e0e0","#8c0052"), pt.size = 1.25, order = T, ncol = 1)
ggsave("Supp_Fig1D_MALATmt.png", plot = p5, device = "png", width = 5, height = 10) 
```

Supp 1E - metaanalysis
Use Zeng 2023 https://doi.org/10.1016/j.stem.2023.04.016 as a reference and find a signature of other germ layers and tissues in our co-culture dataset, focusing specifically on the AxD cluster.
```{r}
data <- read.table("Zeng2023_earlydev/cluster_markers.txt", header = T, sep = ",")
data$celltype <- as.factor(data$celltype)

for (i in levels(data$celltype)){
  goi <- data[data$celltype == i,] %>% arrange(desc(logfoldchanges))
  gene_list <- goi$names
  write(paste0("Zeng2023_earlydev/markers",i,".txt"), x = gene_list)
}

seurat <- readRDS("seurat_all_merged_filtered_mt_rib.rds")

dir <- dir("Zeng2023_earlydev/markers", pattern = ".txt") 
dir.link <- "Zeng2023_earlydev/markers/" 

pdf("Zeng2023_earlydev/reference.pdf", width = 20, height = 10)
for(i in 1:length(dir)){
  gene.set <- as.vector(read.table(paste0(dir.link,dir[i]), header = F)[,1])
  if (length(gene.set) > 0) {
    print(paste("AddModuleScore", dir[i], "running...", sep = " "))
    seurat.scored <- AddModuleScore(seurat, features = list(gene.set) %>% head(n=20), assay = "RNA", name = "reference_score_")
    p <- FeaturePlot(seurat.scored, features = "reference_score_1", 
                     min.cutoff = "q10", max.cutoff = "q90", reduction = "umap", split.by = "Genotype") +
      ggtitle(gsub(".txt", "", dir[i]))
    print(p) 
    rm(seurat.scored)
  }
}
dev.off()
```

Enrichment analysis
```{r}
seurat <- readRDS("seurat_all_merged_filtered_mt_rib.rds")
data <- read.table("Zeng2023_earlydev/cluster_markers_TabS2.txt", header = T, sep = ",", )
data$celltype <- as.factor(data$celltype)
Term2Gene <- data.frame(ID = data$celltype,
                        Gene = data$Gene) 
                        #Gene = data$names)

markers <- read.table("Fig1_markers_precursors_UP.txt", header = F)[,1]

erch <- enricher(gene = markers,
                 universe = seurat@assays[["RNA"]]@counts[rowSums(seurat@assays[["RNA"]]) > 0,] %>% row.names,
                 pAdjustMethod = "fdr",
                 pvalueCutoff = 0.1,
                 qvalueCutoff = 0.2,
                 maxGSSize = 800, 
                 minGSSize = 5,
                 TERM2GENE = Term2Gene)

# barplot with custom enrichment analysis results
custom_bar_plot <- function(res,name,ncat){
  data <- res@result
  data$log.p.adj <- -log10(data$p.adjust) 
  data <- data[order(data$log.p.adj, decreasing = T)[1:ncat],c("Description","p.adjust","log.p.adj")]
  
  enrich.bar <- ggplot(data = data, aes(x = log.p.adj, y = reorder(Description,log.p.adj), fill = p.adjust)) + 
  geom_bar(stat = "identity", width = 0.9) +
  theme_void() +
  theme(axis.text.x = element_text(angle = 0, size = 14, vjust = -0.5, hjust = 0.5), 
        axis.text.y = element_text(size=16, margin = margin(r = 0.2, unit = "cm")),
        axis.line.x = element_line(linewidth = 1, lineend = "square"), 
        axis.line.y = element_blank(),
        axis.title.x = element_text(size = 14, vjust = -2),
        axis.title.y = element_blank(),
        axis.ticks.x = element_line(linewidth = 1, color = "black"),
        axis.ticks.length = unit(0.1, "cm"),
        axis.ticks.y = element_blank(),
        legend.position = "none",
        plot.title = element_text(size = 16, hjust = 0.5, vjust = 2, face = "bold"),
        plot.margin = margin(l = 0.5, r=0.5, b = 0.5, t = 0.5, unit = "cm")) +
  scale_x_continuous(limits = c(0, 25)) +
  coord_capped_cart(bottom = "both") +
   
  ggtitle(paste0("Enrichment analysis", name)) + 
  xlab(expression("-Log"[10]*"(p"[adj]*")")) +
  scale_fill_gradientn(colors = c("#fde725","#5ec962","#31688e"))
  
  return(enrich.bar)
}  

erch@result <- erch@result[erch@result$p.adjust < 0.2,]  
p <- custom_bar_plot(erch, ": Zeng et al. 2023", length(rownames(erch@result)))
p
ggsave(plot = p, 
       filename = "Zeng2023_earlydev/enrichment_Zeng.svg", 
       device = "svg", width = 6, height = 3)
```

# Supplement 2
Supp A-C
```{r}
# A astrocytes
DimPlot(seurat_iA)
brewer.pal(9, "Blues")
display.brewer.pal(9, "Blues")
myCol <- c("#C6DBEF", "#6BAED6", "#2171B5", "#08306B")

p <- DimPlot(seurat_iA, label = FALSE, repel = TRUE, pt.size = 2,
             cols = myCol) +
  theme_void() + 
  NoLegend() + 
  annotate("segment", x = -7, xend = -5.5, y = -6, yend = -6, colour = "black", size=0.75, arrow=arrow(type = "open")) +
  annotate("segment", x = -7, xend = -7, y = -6, yend = -4.75, colour = "black", size=0.75, arrow=arrow(type = "open"))
p 
ggsave("Supp_Fig2A_UMAP_iA.png", plot = p, device = "png", width = 7, height = 7)

q <- DimPlot(seurat_iA, label = FALSE, repel = TRUE, pt.size = 1, split.by = "Sample",
             cols = myCol) +
  theme_void() +
  theme(text = element_text(size = 14))
q 
ggsave("Supp_Fig2A_UMAP_iA_split.png", plot = q, device = "png", width = 10, height = 3)

`ASTRO 1` <- c("HSPB6","ITGA7","S100A2","CRLF1","RBP4")
`ASTRO 2` <- c("GFAP","S100B","COL2A1","FABP7")
`ASTRO 3` <- c("COL1A1","COL3A1","COL1A2","FN1")
`ASTRO 4` <- c("TK1","NUSAP1","MKI67","CDK1")

p.dot <- DotPlot(seurat_iA, features = c(`ASTRO 1`,`ASTRO 2`,`ASTRO 3`,`ASTRO 4`),
                 cols = c("#f7f7f7","#2171B5")) +
  theme(axis.title = element_text(size = 10),
        axis.title.y = element_text(vjust = 2),
        axis.text.x = element_text(angle=60, vjust = 1, hjust =1),
        legend.title = element_text(size = 8), legend.text = element_text(size = 8), 
        legend.key.size = unit(0.35, "cm"), legend.text.align = 1, legend.title.align = 0, 
        plot.margin = margin(r = 1, l = 0.25, t = 0.25, b = 0.25, unit = "cm"),
        legend.margin = margin(t = 0, r = 0, l = 0, b = -0.25, unit = "cm"),
        plot.title = element_text(size = 14,face = "plain", hjust = 0.5)) +
  guides(colour = guide_colorbar(order = 1, title = "Scaled Average\nExpression"), 
         size = guide_legend(order = 2, title = "Percent\nExpressed")) +
  xlab("Genes") + ylab("Cluster Identity") +
  ggtitle("Astrocyte cluster markers")
p.dot
ggsave("Supp_Fig2A_marker_iA.png", device = "png", height = 3, width = 7, bg = "#FFFFFF")

# B neurons
DimPlot(seurat_iN)
brewer.pal(9, "Reds")
display.brewer.pal(9, "Reds")
myCol <- c("#FCBBA1", "#FC9272", "#EF3B2C", "#A50F15")

p <- DimPlot(seurat_iN, label = F, repel = TRUE, pt.size = 2,
             cols = myCol, label.size = 5) +
  theme_void() + 
  NoLegend() + 
  annotate("segment", x = -6, xend = -4.75, y = -5, yend = -5, colour = "black", size=0.75, arrow=arrow(type = "open")) +
  annotate("segment", x = -6, xend = -6, y = -5, yend = -4, colour = "black", size=0.75, arrow=arrow(type = "open"))
p 
ggsave("Supp_Fig2B_UMAP_iN.png", plot = p, device = "png", width = 7, height = 7)

q <- DimPlot(seurat_iN, label = FALSE, repel = TRUE, pt.size = 1, split.by = "Sample",
             cols = myCol) +
  theme_void() +
  theme(text = element_text(size = 14))
q 
ggsave("Supp_Fig2B_UMAP_iN_split.png", plot = q, device = "png", width = 10, height = 3)

`NEURO 1` <- c("FTL","SERF2","IGFBP2","TNFRSF12A","VIM","S100A11")
`NEURO 2` <- c("SCG2","NNAT","PCSK1","SYT4","BDNF","STC1")
`NEURO 3` <- c("PTH2","UCN","SCG5","TRH")
`NEURO 4` <- c("PRPH","ISL1","PHOX2B","PCP4","IFI16","LINC00682")

p.dot <- DotPlot(seurat_iN, features = c(`NEURO 1`,`NEURO 2`,`NEURO 3`,`NEURO 4`),
                 cols = c("#f7f7f7","#A50F15")) +
  theme(axis.title = element_text(size = 10),
        axis.title.y = element_text(vjust = 2),
        axis.text.x = element_text(angle=60, vjust = 1, hjust =1),
        legend.title = element_text(size = 8), legend.text = element_text(size = 8), 
        legend.key.size = unit(0.35, "cm"), legend.text.align = 1, legend.title.align = 0, 
        plot.margin = margin(r = 1, l = 0.25, t = 0.25, b = 0.25, unit = "cm"),
        legend.margin = margin(t = 0, r = 0, l = 0, b = -0.25, unit = "cm"),
        plot.title = element_text(size = 14,face = "plain", hjust = 0.5)) +
  guides(colour = guide_colorbar(order = 1, title = "Scaled Average\nExpression"), 
         size = guide_legend(order = 2, title = "Percent\nExpressed")) +
  xlab("Genes") + ylab("Cluster Identity") +
  ggtitle("Neuron cluster markers")
p.dot
ggsave("Supp_Fig2B_marker_iN.png", device = "png", height = 3, width = 7, bg = "#FFFFFF")

# C AxD cluster
DimPlot(seurat_iPSC)
brewer.pal(9, "Greens")
display.brewer.pal(9, "Greens")
myCol <- c("#C7E9C0","#74C476","#238B45","#00441B")
levels(seurat_iPSC) <- as.factor(c("EPCAM/CDH1", "PRE-ASTRO","PRE-NEURO","DCX/STMN1"))

p <- DimPlot(seurat_iPSC, label = FALSE, repel = TRUE, pt.size = 2,
             cols = myCol) +
  theme_void() + 
  NoLegend() + 
  annotate("segment", x = -7, xend = -5.5, y = -5, yend = -5, colour = "black", size=0.75, arrow=arrow(type = "open")) +
  annotate("segment", x = -7, xend = -7, y = -5, yend = -3.75, colour = "black", size=0.75, arrow=arrow(type = "open"))
p 
ggsave("Supp_Fig2C_UMAP_iPSC.png", plot = p, device = "png", width = 7, height = 7)

q <- DimPlot(seurat_iPSC, label = FALSE, repel = TRUE, pt.size = 1, split.by = "Sample",
             cols = myCol) +
  theme_void() +
  theme(text = element_text(size = 14))
q 
ggsave("Supp_Fig2C_UMAP_iPSC_split.png", plot = q, device = "png", width = 10, height = 3)

#marker dotplot
levels(seurat_iPSC) <- as.factor(c("EPCAM/CDH1", "PRE-ASTRO","PRE-NEURO","DCX/STMN1"))
p.dot <- DotPlot(seurat_iPSC, features = c("CDH1","EPCAM","CLDN7","ITGA2","ITGA3",
                                           "CHI3L2","MGP","S100A1","S100B","COL2A1",
                                           "VIM",
                                           "GPX1","VAMP8","GSTP1","CYTOR","PDLIM1",
                                           "RTN1","STMN1","STMN2","DCX","GAP43","VGF"),
                 cols = c("#f7f7f7","#238B45")) +
  theme(axis.title = element_text(size = 10),
        axis.title.y = element_text(vjust = 2),
        axis.text.x = element_text(angle=60, vjust = 1, hjust =1),
        legend.title = element_text(size = 8), legend.text = element_text(size = 8), 
        legend.key.size = unit(0.35, "cm"), legend.text.align = 1, legend.title.align = 0, 
        plot.margin = margin(r = 1, l = 0.25, t = 0.25, b = 0.25, unit = "cm"),
        legend.margin = margin(t = 0, r = 0, l = 0, b = -0.25, unit = "cm"),
        plot.title = element_text(size = 14,face = "plain", hjust = 0.5)) +
  guides(colour = guide_colorbar(order = 1, title = "Scaled Average\nExpression"), 
         size = guide_legend(order = 2, title = "Percent\nExpressed")) +
  xlab("Genes") + ylab("Cluster Identity") +
  ggtitle("Precursor cluster markers")
p.dot
ggsave(plot = p.dot, filename = "Supp_Fig2C_marker_iPSC.png", device = "png", height = 3, width = 7.5, bg = "#FFFFFF")
```

Supp 2D
```{r}
reference.gene.set <- c("GFAP","HSPB1","VIM","CRYAB","PLEC","SYNM","CCND2","SQSTM1") #Rosenthal fibers 
seurat.object <- readRDS("iPSC_iA/seurat_iA_filtered_mt_rib.rds")
seurat.object@active.assay <- "RNA"
seurat.object.subset <- subset(seurat.object, subset = GFAP > 0, idents = "ASTRO_2")

df <- as.data.frame(AverageExpression(seurat.object.subset, features = reference.gene.set, assays = "RNA", group.by = "Sample", slot = "scale.data"))

colnames(df) <- c("CTRL AxD/AxD", "CTRL AxD/C", "CTRL C/C",
                                  "OGD AxD/AxD", "OGD AxD/C", "OGD C/C")
df <- df[,c(3,2,1,6,5,4)]
mat <- as.matrix(df)

svg("Supp_Fig2D_heatmap_RFs.svg", width = 4, height = 5)
  draw(Heatmap(mat, col = circlize::colorRamp2(c(-1,0,3),c("blue","#eeeeee","red")),
               column_title = "Rosenthal fiber genes - ASTRO 2 GFAP+",
               column_title_gp = gpar(fontsize = 9, fontface = "bold"), 
               cluster_rows = T, cluster_columns = F, show_row_dend = F, 
               heatmap_legend_param = list(title = expression("Scaled\nexpression"))))
dev.off()
```

Supp 2E
```{r}
seurat.object <- readRDS("seurat_all_merged_filtered_mt_rib.rds")
seurat.object@active.assay <- "RNA"
p <- FeaturePlot(seurat.object, features = c("SOX9","NFIB","NEUROG2"), pt.size = 0.5, min.cutoff = "q10", max.cutoff = "q90", ncol = 1, cols = c("#e0e0e0","#8c0052"), order = T) & NoLegend() 
ggsave("Supp_Fig2E_precursor_origin.png", plot = p, device = "png", width = 3, height = 8)
```

# Supplement 3
Supp 3A
```{r}
# cellchat 1 = wt, cellchat 2 = AxD 
pathways.show <- c("FGF","TGFb","LIFR","ncWNT","NOTCH","EGF")
plots <- lapply(pathways.show, function(x) netAnalysis_contribution(cellchat1, signaling = x, title = paste0("L-R pairs in ",x)))
ggsave("Supp_Fig3A_wt_L-Rpairs.png", plot = egg::ggarrange(plots = plots), width = 7, height = 4)

# Supp. 3B
pathways.show <- c("ncWNT","NOTCH","EGF","RELN")
plots <- lapply(pathways.show, function(x) netAnalysis_contribution(cellchat2, signaling = x, title = paste0("L-R pairs in ",x)))
ggsave("Supp_Fig3B_wa_L-Rpairs.png", plot = egg::ggarrange(plots = plots), width = 7, height = 4)
```

Supp 3C
```{r}
gg6 <- rankNet(cellchat.merged, mode = "comparison", stacked = T, do.stat = TRUE, color.use = c("#7AA3CD","#D17480"))
ggsave("Supp_Fig3C_CellChat_infflow.png",plot = gg6 + theme(legend.position = "none"), device = "png", width = 3, height = 5.5)
```

Supp 3D
```{r}
pathways.show <- c("FGF","TGFb","LIFR","ncWNT","NOTCH","EGF","RELN")
ht1 = netAnalysis_signalingRole_heatmap(cellchat2, pattern = "outgoing", signaling = pathways.show, title = "AxD outgoing", width = 6, height = 6)

svg("Supp_Fig3D_CellChat_outgoing.svg")
ComplexHeatmap::draw(ht1)
dev.off()

ht3 = netAnalysis_signalingRole_heatmap(cellchat2, pattern = "incoming", signaling = pathways.show, title = "AxD incoming", width = 6, height = 6, color.heatmap = "GnBu")

svg("Supp_Fig3D_CellChat_incoming.svg")
ComplexHeatmap::draw(ht3)
dev.off()
```

Supp 4A
```{r}
DE_iA <- readRDS("iPSC_iA/DEA_mt_rib/DE_iA.rds")[1:2]
names(DE_iA) <- c("astroC/neuroC - astroAxD/neuroAxD",
                  "astroC/neuroC - astroAxD/neuroC"
                  )
gene.lists <- lapply(DE_iA[c("astroC/neuroC - astroAxD/neuroAxD","astroC/neuroC - astroAxD/neuroC")], 
                    function(x) x = x[(x$avg_log2FC > 0.65) & x$p_val_adj < 0.05,])
gene.lists <- lapply(DE_iA[c("astroC/neuroC - astroAxD/neuroAxD","astroC/neuroC - astroAxD/neuroC")], 
                     function(x) x = x[(x$avg_log2FC < -0.65) & x$p_val_adj < 0.05,])
venn <- lapply(gene.lists, function(x) rownames(x))
#venn.plot <- ggvenn(venn, fill_color = c("#EB7321","#EB7321"), fill_alpha = 0.4,
venn.plot <- ggvenn(venn, fill_color = c("#5A7BC6","#5A7BC6"), fill_alpha = 0.4,
                    set_name_size = 4.5, show_percentage = F, text_size = 7) + 
  NoLegend()
ggsave("Supp_Fig4A_Venn_up.png",plot = venn.plot, 
       device = "png", bg = "white",
       width = 6, height = 4)
ggsave("Supp_Fig4A_Venn_down.png",plot = venn.plot, 
       device = "png", bg = "white",
       width = 6, height = 4)
```

Supp 4B
```{r}
DE_tests <- readRDS("iPSC_iA/DEA_mt_rib/DE_iA.rds")
DE_tests <- readRDS("iN/DEA_mt_rib/DE_iN.rds")
res <- DE_tests[["Ctrl_w_a-Ctrl_AxD"]]

keyvals.colour <- ifelse(
  res$avg_log2FC < -0.65 & res$p_val_adj < 0.05, '#5A7BC6',
  ifelse(res$avg_log2FC > 0.65 & res$p_val_adj < 0.05, '#EB7321',
         '#E7E7E7')) 
names(keyvals.colour)[keyvals.colour == '#5A7BC6'] <- 'downregulated'
names(keyvals.colour)[keyvals.colour == '#EB7321'] <- 'upregulated'
names(keyvals.colour)[keyvals.colour == '#E7E7E7'] <- 'insignificant'

top10up <- res[res$avg_log2FC > 0.65 & res$p_val_adj < 0.05,]
top10up <- rownames(head(top10up[order(top10up$p_val_adj, decreasing = F),],10))
top10down <- res[res$avg_log2FC < -0.65 & res$p_val_adj < 0.05,]
top10down <- rownames(head(top10down[order(top10down$p_val_adj, decreasing = F),],10))

p <- EnhancedVolcano(toptable = res,
                     lab = rownames(res),
                     selectLab = c(top10up,top10down,"GFAP"),
                     x = 'avg_log2FC',
                     y = 'p_val_adj',
                     title = "iN: astroAxD/neuroC - astroAxD/neuroAxD",
                     pCutoff = 0.05, 
                     colAlpha = 0.7,
                     colCustom = keyvals.colour,
                     FCcutoff = 0.65, 
                     labSize = 4.5,
                     pointSize = 3,
                     drawConnectors = T,
                     maxoverlapsConnectors = Inf,
                     min.segment.length = 0.1,
                     cutoffLineType = "blank",
                     arrowheads = F,
                     caption = "|log2FC| > 0.65 and p-adj < 0.05") +
  theme_light() +
  theme(plot.subtitle = element_blank(), 
        plot.caption = element_text(size = 10), 
        plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        axis.line = element_line(colour = "black", linewidth = 0.7, linetype = "solid"), 
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        axis.ticks = element_line(linewidth = 0.7, colour = "black"),
        axis.ticks.length = unit(0.2, "cm"),
        legend.position = "none") +
  xlab(expression("Log"[2]*"FC")) +
  guides(color = guide_legend(override.aes = list(size = 4)))
p
ggsave(plot = p, "Supp_Fig4B_volcano_iA.png", device = "png", width = 6, height = 5)
ggsave(plot = p, "Supp_Fig4B_volcano_iN.png", device = "png", width = 6, height = 5)
```

Supp 4C
```{r}
gse <- readRDS("iA_GSEA.rds")
gse@result <- gse@result[order(gse@result$p.adjust),]
gse@result$Core_count <- sapply(strsplit(gse@result$core_enrichment,split = "/"), FUN = length)
gse@result <- gse@result[gse@result$Core_count > 3,]
p <- dotplot(gse, showCategory = 5, label_format = 50, split=".sign", orderBy = "GeneRatio") + facet_grid(.~.sign) +
  scale_color_gradient(low = "#8c0052",high = "#e0e0e0") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 16, colour = "black"),
        axis.text.x = element_text(size = 12),
        plot.title = element_text(size = 16, face = "bold"),
        strip.text.x = element_text(size = 16, face = "bold"),
        strip.background = element_rect(linetype = "blank",colour = "white",fill=c("#e7e7e7"), size=1.5)) + 
  ggtitle("iA: GO Terms Enrichment")
p
ggsave(plot = p, filename = "Supp_Fig4C_iA_GSEA.png", device = "png", bg = "white", width = 10, height = 5)
```

Supp 4D
```{r}
DE_tests_list <- list(Precursors = readRDS("iPSC_iA/DEA_mt_rib/DE_iPSC.rds")[["Ctrl_w_a-OGD_w_a"]],
                      iA = readRDS("iPSC_iA/DEA_mt_rib/DE_iA.rds")[["Ctrl_w_a-OGD_w_a"]],
                      iN = DE_tests <- readRDS("iN/DEA_mt_rib/DE_iN.rds")[["Ctrl_w_a-OGD_w_a"]])

select_labels <- list(iPSC = c("C4orf48","PCSK1N","SNHG9","SNHG25","SEC61G","NDUFA3","TMSB10","TCEB2","NDUFB1","UQCRQ","USMG5","UBL5","NAA38","MT2A"),
                      iA = c("PCSK1N","HES4","C4orf48","KRT19","SNHG9","VGF","STMN1","WFDC2"),
                      iN = c("FN1","THBS1","SPARC","PTH2"))

volcano_custom <- function(title, labelGenes, DE_tests, plotFileName){
  res <- DE_tests 
  keyvals.colour <- ifelse(
  res$avg_log2FC < -0.65 & res$p_val_adj < 0.05, '#5A7BC6',
  ifelse(res$avg_log2FC > 0.65 & res$p_val_adj < 0.05, '#EB7321',
         '#E7E7E7')) 
  names(keyvals.colour)[keyvals.colour == '#5A7BC6'] <- 'downregulated'
  names(keyvals.colour)[keyvals.colour == '#EB7321'] <- 'upregulated'
  names(keyvals.colour)[keyvals.colour == '#E7E7E7'] <- 'insignificant'
  
  p <- EnhancedVolcano(toptable = res,
                     lab = rownames(res),
                     selectLab = labelGenes,
                     x = 'avg_log2FC',
                     y = 'p_val_adj',
                     axisLabSize = 8,
                     titleLabSize = 10,
                     title = title,
                     pCutoff = 0.05, 
                     colAlpha = 0.7,
                     colCustom = keyvals.colour,
                     subtitleLabSize = 9,
                     FCcutoff = 0.65, 
                     labSize = 4.5,
                     pointSize = 3,
                     drawConnectors = T,
                     maxoverlapsConnectors = Inf,
                     min.segment.length = 0.1,
                     arrowheads = FALSE,
                     boxedLabels = F,
                     cutoffLineType = "blank",
                     caption = "|log2FC| > 0.65 and p-adj < 0.05") +
  theme_light() +
  theme(plot.subtitle = element_blank(), 
        plot.caption = element_text(size = 10), 
        plot.title = element_text(hjust = 0.5, vjust = 1, size = 13, face = "bold"),
        panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        axis.line = element_line(colour = "black", linewidth = 0.7, linetype = "solid"), 
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        axis.ticks = element_line(linewidth = 0.7, colour = "black"),
        axis.ticks.length = unit(0.2, "cm"),
        legend.position = "none") +
  xlab(expression("Log"[2]*"FC")) +
  guides(color = guide_legend(override.aes = list(size = 4)))
  
  ggsave(plot = p, filename = plotFileName, device = "png", width = 6, height = 5)
  return(p)
}

lapply(seq_along(DE_tests_list), function(x) volcano_custom(DE_tests_list[[x]], 
                                                            labelGenes = select_labels[[x]], 
                                                            title = paste0(names(DE_tests_list)[x],": CTRL astroAxD/neuroC - OGD astroAxD/neuroC "), 
                                                            plotFileName = paste0("Supp_Fig4D_",names(DE_tests_list)[x],"_volcano.png")))
```

Supp 4E
```{r}
DE.object <- readRDS("iPSC_iA/DEA_mt_rib/DE_iPSC.rds")

gene.list <- subfunction1(DE.object[["Ctrl_w_a-OGD_w_a"]]) # create ranked gene list
  
gse <- gseGO(geneList=gene.list, 
            ont ="ALL", 
            keyType = "ALIAS", 
            #nPermSimple = 10000, 
            minGSSize = 3, 
            maxGSSize = 800, 
            pvalueCutoff = 0.1, 
            verbose = TRUE, 
            OrgDb = org.Hs.eg.db, 
            pAdjustMethod = "BH",
            by = "fgsea")
gse <- simplify(gse)

gse@result <- gse@result[order(gse@result$p.adjust),]
gse@result$Core_count <- sapply(strsplit(gse@result$core_enrichment,split = "/"), FUN = length)
gse@result <- gse@result[gse@result$Core_count > 3,]
p <- dotplot(gse, showCategory = 5, label_format = 50, split=".sign", orderBy = "GeneRatio") + facet_grid(.~.sign) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 15, colour = "black"),
        axis.text.x = element_text(size = 12),
        plot.title = element_text(size = 16, face = "bold"),
        strip.text.x = element_text(size = 16, face = "bold"),
        strip.background = element_rect(linetype = "blank",colour = "white",fill=c("#e7e7e7"), size=1.5)) + 
  ggtitle("Precursors OGD: GO Terms Enrichment") +
  scale_color_gradient(low = "#8c0052",high = "#e0e0e0")
p
ggsave(plot = p, filename = "Supp_Fig4E_precursors_GSEA_dotplot.png", device = "png", bg = "white", width = 9.5, height = 5)
```


