---
title: "CellChat"
author: "Sarka"
date: "2/16/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

load libraries
```{r}
library(NMF)
library(ggalluvial)
library(Seurat)
library(CellChat)
library(ComplexHeatmap)
library(Seurat)
setwd("CellChat/OGD/")
```

prepare data from Seurat
```{r}
seurat <- readRDS("seurat_all_merged_filtered_mt_rib.rds")
sample.name <- "Ctrl_w_a"
seurat.subset <- subset(seurat, subset = Sample == sample.name)
seurat.subset@active.assay <- "RNA"
seurat.subset <- RenameIdents(seurat.subset, 
                              "EPCAM/CDH1" = "Precursors",
                              "PRE-ASTRO" = "Precursors",
                              "PRE-NEURO" = "Precursors",
                              "DCX/STMN1" = "Precursors")
seurat.subset <- subset(seurat.subset, idents = c("Precursors","ASTRO_1","ASTRO_2","ASTRO_3","NEURO_1","NEURO_2","NEURO_3","NEURO_4"))
all.genes <- rownames(seurat.subset)
seurat.subset <- NormalizeData(seurat.subset, normalization.method = "LogNormalize", scale.factor = 10000, assay = "RNA")
seurat.subset <- ScaleData(seurat.subset, features = all.genes, assay = "RNA")

cellchat <- createCellChat(object = seurat.subset, group.by = "ident")
saveRDS(cellchat, paste0(sample.name,"_cellchat.rds"))
```

load data
```{r}
cellchat <- readRDS(paste0(sample.name,"_cellchat.rds"))
```

add DB
```{r}
CellChatDB <- CellChatDB.human 
showDatabaseCategory(CellChatDB)
CellChatDB.use <- CellChatDB
cellchat@DB <- CellChatDB.use
```

Processing
```{r}
cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database
  future::plan("multiprocess", workers = 2) # do parallel
  cellchat <- identifyOverExpressedGenes(cellchat)
  cellchat<- identifyOverExpressedInteractions(cellchat)
   cellchat <- computeCommunProb(cellchat, population.size = TRUE)
  # Filter out the cell-cell communication if there are only few number of cells in certain cell groups
  cellchat <- filterCommunication(cellchat, min.cells = 10)
  #df.net <- subsetCommunication(cellchat_list[[i]],slot.name = "net")
  cellchat <- computeCommunProbPathway(cellchat)
  cellchat <- aggregateNet(cellchat)
  cellchat <- netAnalysis_computeCentrality(cellchat)
saveRDS(cellchat,paste0(sample.name,"_cellchat_processed.rds"))
```

Loading processed cellchat dataset
```{r}
cellchat <- readRDS(paste0(sample.name,"_cellchat_processed.rds"))
```

Interaction numbers and strength
Note: Most of the signaling happens between astrocytes (all subtypes) - note the autocrine signaling too.
```{r}
par(mfrow = c(1,2), xpd=TRUE)
groupSize <- as.numeric(table(cellchat@idents))
cir1 <- netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
cir2 <- netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength")
```

Heatmap for all sinaling and cell types
```{r}
pathway.show <- cellchat@netP$pathways
ht1 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing", signaling = pathway.show, width = 8, height = 15)
ht2 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming", signaling = pathway.show, width = 8, height = 15)
ht1+ht2
```

Signaling patterns
Note: Incoming signal - Astro_3 and Neuro_4 similar signaling. Outgoing - Precursors and Neuro_4 similar.
```{r}
x <-selectK(cellchat, pattern = "outgoing")
y <-selectK(cellchat, pattern = "incoming")
nPatternsx <- 5
nPatternsy <- 5
cellchat <- identifyCommunicationPatterns(cellchat, pattern = "outgoing", k = nPatternsx)
cellchat <- identifyCommunicationPatterns(cellchat, pattern = "incoming", k = nPatternsy)
par(mfrow=c(1,1))
ht3 <- netAnalysis_river(cellchat, pattern = "incoming")
ht4 <- netAnalysis_river(cellchat, pattern = "outgoing")
```

Dominant senders and receivers - cell types overall
Note: 3 astro groups -> main senders and receivers. Precursors and Neuro_2 receive more than send.
```{r}
gg1 <- netAnalysis_signalingRole_scatter(cellchat)
gg1
```

Save to rds
```{r}
saveRDS(cellchat, paste0(sample.name,"_cellchat_processed.rds"))
```

```{r}
pdf(paste0(sample.name,"_result.pdf"), height = 10, width = 7)
print(cir1)
print(cir2)
print(ht1)
print(ht2)
print(gg1)
dev.off()

# for (i in cellchat@netP$pathways) {
#   png(file =paste0("Pathways_circle_plots/",i,"_Ctrl_wt",".png"), width = 1024, height = 768)
#   netVisual_aggregate(cellchat, signaling = i, layout = "circle")
#   dev.off()
# }
```

Compare datasets
Prepare data
```{r}
cellchat1 <- readRDS("CellChat/CTRL/Ctrl_wt_cellchat_processed.rds")
cellchat2 <- readRDS("CellChat/CTRL/Ctrl_w_a_cellchat_processed.rds")
sample.names <- c("Ctrl_wt","Ctrl_w_a")

cellchat.merged <- mergeCellChat(list(cellchat1,cellchat2), add.names = sample.names)
```

Compare total number of interactions and their strength
```{r}
gg1 <- compareInteractions(cellchat.merged, show.legend = F, group = c(1,2))
gg2 <- compareInteractions(cellchat.merged, show.legend = F, group = c(1,2), measure = "weight")
gg1 + gg2

par(mfrow = c(1,2), xpd=TRUE)
cir1 <- netVisual_diffInteraction(cellchat.merged, weight.scale = T)
cir2 <- netVisual_diffInteraction(cellchat.merged, weight.scale = T, measure = "weight")

gg3 <- netVisual_heatmap(cellchat.merged)
gg4 <- netVisual_heatmap(cellchat.merged, measure = "weight")
gg3 + gg4
```

Compare major sources and pathways
```{r}
object.list <- list(cellchat1,cellchat2); names(object.list) <- sample.names
num.link <- sapply(object.list, function(x) {rowSums(x@net$count) + colSums(x@net$count)-diag(x@net$count)})
weight.MinMax <- c(min(num.link), max(num.link)) # control the dot size in the different datasets
gg5 <- list()
for (i in 1:length(object.list)) {
  gg5[[i]] <- netAnalysis_signalingRole_scatter(object.list[[i]], title = names(object.list)[i], weight.MinMax = weight.MinMax)
}

patchwork::wrap_plots(plots = gg5)
ggsave("wt_major_sources.png",plot = patchwork::wrap_plots(plots = gg5), device = "png", width = 7, height = 5)

# gg0 <- netAnalysis_signalingChanges_scatter(cellchat.merged, idents.use = "Precursors")
# 
# # on a single cellchat object
# pathways.show <- c("ncWNT","NOTCH","TGFb")
# netAnalysis_contribution(cellchat2, signaling = pathways.show)
```

Compare information flow
```{r}
gg6 <- rankNet(cellchat.merged, mode = "comparison", stacked = T, do.stat = TRUE)
gg7 <- rankNet(cellchat.merged, mode = "comparison", stacked = F, do.stat = TRUE)
gg6 + gg7

ggsave("wt_information_flow.png",plot = gg7, device = "png", width = 4, height = 5)
```

Outgoing and incoming signaling
```{r}
 
# combining all the identified signaling pathways from different datasets 
pathway.union <- union(object.list[[i]]@netP$pathways, object.list[[i+1]]@netP$pathways)

ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "outgoing", signaling = pathway.union, title = names(object.list)[i], width = 6, height = 12)
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "outgoing", signaling = pathway.union, title = names(object.list)[i+1], width = 6, height = 12)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))

ht3 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "incoming", signaling = pathway.union, title = names(object.list)[i], width = 6, height = 12, color.heatmap = "GnBu")
ht4 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "incoming", signaling = pathway.union, title = names(object.list)[i+1], width = 6, height = 12, color.heatmap = "GnBu")
draw(ht3 + ht4, ht_gap = unit(0.5, "cm"))

ht5 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "all", signaling = pathway.union, title = names(object.list)[i], width = 6, height = 12, color.heatmap = "OrRd")
ht6 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "all", signaling = pathway.union, title = names(object.list)[i+1], width = 6, height = 12, color.heatmap = "OrRd")
draw(ht5 + ht6, ht_gap = unit(0.5, "cm"))
```

Identify dysregulated ligand-receptor pairs
```{r}
# gg8 <- netVisual_bubble(cellchat.merged, sources.use = 4, targets.use = c(5:11),  comparison = c(1, 2), max.dataset = 2, title.name = paste0("Increased signaling in ",sample.names[2]), angle.x = 45, remove.isolate = T)
# #> Comparing communications on a merged object
# gg9 <- netVisual_bubble(cellchat.merged, sources.use = 4, targets.use = c(5:11),  comparison = c(1, 2), max.dataset = 1, title.name = paste0("Decreased signaling in ",sample.names[2]), angle.x = 45, remove.isolate = T)
# #> Comparing communications on a merged object
# gg8 + gg9
```

Visually compare in circle plots, pathway needs to be present in both datasets
```{r}
data <- rankNet(cellchat.merged, mode = "comparison", stacked = F, do.stat = TRUE, return.data = T)
new.dir <- paste0(paste(names(object.list),collapse='_'),"_plots")
dir.create(new.dir)
setwd(new.dir)

for (pw in levels(as.factor(data[["signaling.contribution"]]$name))){
  pathways.show <- pw
  png(paste0(pw,".png"), width = 1024, height = 768)
  if (pathways.show %in% object.list[[1]]@netP[["pathways"]] & pathways.show %in% object.list[[2]]@netP[["pathways"]]){
    weight.max <- getMaxWeight(object.list, slot.name = c("netP"), attribute = pathways.show) # control the edge weights across different datasets
    par(mfrow = c(1,2), xpd=TRUE)
    for (i in 1:length(object.list)) {
       netVisual_aggregate(object.list[[i]], signaling = pathways.show, layout = "circle", 
                            edge.weight.max = weight.max[1], edge.width.max = 10, 
                            signaling.name = paste(pathways.show, names(object.list)[i]))
    }
  } else {
    if (pathways.show %in% object.list[[1]]@netP[["pathways"]]){
      vertex.receiver = seq(2,4)
      print(netVisual_aggregate(object.list[[1]], signaling = pathways.show, 
                          layout = "hierarchy",vertex.receiver = vertex.receiver))
    } else {
      vertex.receiver = seq(2,4)
        print(netVisual_aggregate(object.list[[2]], signaling = pathways.show, 
                            layout = "hierarchy",vertex.receiver = vertex.receiver))
    }
  }
  dev.off()
}
```

Save plots
```{r}
pdf(paste0(paste(names(object.list),collapse='_'),"_results.pdf"), height = 10, width = 7)
print(gg1)
print(gg2)
print(cir1)
print(cir2)
print(gg3)
print(gg4)
print(gg5)
print(ht1)
print(ht2)
print(ht3)
print(ht4)
print(ht5)
print(ht6)
print(gg6)
print(gg7)
# print(gg8)
# print(gg9)
dev.off()
```

Interesting significant pathways
```{r}
pathways.show <- c("RELN")

png("RELN_pathway_net.png",res = 300, height = 1000, width = 1500)
netAnalysis_signalingRole_network(cellchat, signaling = pathways.show, width = 8, height = 4, font.size = 10,)
dev.off()

#netAnalysis_contribution(cellchat, signaling = pathways.show)

png("RELN_pathway_exp.png",res = 300, height = 1000, width = 1500)
plotGeneExpression(cellchat, signaling = pathways.show, colors.ggplot = T) 
dev.off()

# vertex.receiver = seq(2,4)
# netVisual_aggregate(cellchat, signaling = pathways.show, layout = "hierarchy",vertex.receiver = vertex.receiver)

png("RELN_pathway_circ.png",res = 250, height = 1000, width = 1000, )
par(mfrow = c(1,1))
netVisual_aggregate(cellchat, signaling = pathways.show, layout = "circle")
dev.off()

# pairLR.use <- extractEnrichedLR(cellchat, signaling = pathways.show)
# netVisual_bubble(cellchat, sources.use = "NEURO_2", pairLR.use = pairLR.use, remove.isolate = TRUE) + 
#   theme(axis.text.x = element_text(angle = 45, vjust = 1))
```

wt-AxD interesting pathways
```{r}
cellchat1 <- readRDS("CellChat/Ctrl_wt_cellchat_processed.rds")
cellchat2 <- readRDS("CellChat/Ctrl_AxD_cellchat_processed.rds")
sample.names <- c("Ctrl_wt","Ctrl_AxD")

object.list <- list(cellchat1,cellchat2); names(object.list) <- sample.names

#up in AxD
#pathways.show <- c("NOTCH","ACTIVIN","JAM","ncWNT","DESMOSOME","LIFR","TGFb","EGF",
                   #"EPHA","ADGRE5","SOMATOSTATIN","NGL","SEMA7","NRG")
#up in wt
pathways.show <- c("APP","IL6","CHAD","ENHO","CXCL","IGF","NECTIN","CALCR","PROS","GAS","FGF")

i <- 1
ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "outgoing", signaling = pathways.show, title = names(object.list)[i], width = 6, height = 12)
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "outgoing", signaling = pathways.show, title = names(object.list)[i+1], width = 6, height = 12)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))

png("CellChat/wt-AxD_outgoing_wtup.png", width = 2400, height = 2400, res = 300)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))
dev.off()

ht3 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "incoming", signaling = pathways.show, title = names(object.list)[i], width = 6, height = 12, color.heatmap = "GnBu")
ht4 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "incoming", signaling = pathways.show, title = names(object.list)[i+1], width = 6, height = 12, color.heatmap = "GnBu")
draw(ht3 + ht4, ht_gap = unit(0.5, "cm"))

png("CellChat/wt-AxD_incoming_wtup.png", width = 2400, height = 2400, res = 300)
draw(ht3 + ht4, ht_gap = unit(0.5, "cm"))
dev.off()
```

w_a-AxD interesting (unique) pathways
```{r}
cellchat1 <- readRDS("CellChat/Ctrl_w_a_cellchat_processed.rds")
cellchat2 <- readRDS("CellChat/Ctrl_AxD_cellchat_processed.rds")
sample.names <- c("Ctrl_w_a","Ctrl_AxD")

object.list <- list(cellchat1,cellchat2); names(object.list) <- sample.names

pathways.show <- c("RELN","NEGR","NCAM","MK","CD99","CDH","TENASCIN","LAMININ","NECTIN","SEMA3","EPHB","VISFATIN","GDF","MIF","CALCR","FGF","CD46")

i <- 1
ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "outgoing", signaling = pathways.show, title = names(object.list)[i], width = 6, height = 12)
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "outgoing", signaling = pathways.show, title = names(object.list)[i+1], width = 6, height = 12)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))

png("CellChat/w_a-AxD_outgoing.png", width = 2400, height = 2400, res = 300)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))
dev.off()

ht3 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "incoming", signaling = pathways.show, title = names(object.list)[i], width = 6, height = 12, color.heatmap = "GnBu")
ht4 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "incoming", signaling = pathways.show, title = names(object.list)[i+1], width = 6, height = 12, color.heatmap = "GnBu")
draw(ht3 + ht4, ht_gap = unit(0.5, "cm"))

png("CellChat/w_a-AxD_incoming.png", width = 2400, height = 2400, res = 300)
draw(ht3 + ht4, ht_gap = unit(0.5, "cm"))
dev.off()
```

wt-AxD individual pathways
```{r}
cellchat1 <- readRDS("CellChat/Ctrl_wt_cellchat_processed.rds")
cellchat2 <- readRDS("CellChat/Ctrl_AxD_cellchat_processed.rds")
pathways.show <- c("NOTCH","ncWNT","LIFR","TGFb","EGF")

for (pathway in pathways.show){
  png(paste0("CellChat/AxD_",pathway,"_circ.png"),res = 250, height = 1000, width = 1000)
    par(mfrow = c(1,1))
    netVisual_aggregate(cellchat2, signaling = pathway, layout = "circle")
  dev.off()
  
  ggsave(paste0("CellChat/AxD_",pathway,"_L-Rpairs.png"), plot = netAnalysis_contribution(cellchat2, signaling = pathway))
}

pathways.show <- c("PROS","GAS","FGF")
object.list <- list(cellchat1,cellchat2); names(object.list) <- c("Ctrl_wt","Ctrl_AxD")
for (pathway in pathways.show){
  weight.max <- getMaxWeight(object.list, slot.name = c("netP"), attribute = pathway) # control the edge weights across different datasets
  png(paste0("CellChat/wt-AxD_",pathway,"_circ.png"),res = 250, height = 1000, width = 2000)
  par(mfrow = c(1,2), xpd=TRUE)
  for (i in 1:length(object.list)) {
    netVisual_aggregate(object.list[[i]], signaling = pathway, layout = "circle",
                        edge.weight.max = weight.max[1], edge.width.max = 10, 
                        signaling.name = paste(pathway, names(object.list)[i]))
  }  
  dev.off()
  ggsave(paste0("CellChat/wt-AxD_",pathway,"_L-Rpairs.png"), plot = netAnalysis_contribution(object.list[[2]], signaling = pathway))
}  
```

Identify signaling groups based on their functional similarity
```{r}
cellchat.merged <- computeNetSimilarityPairwise(cellchat.merged, type = "functional")
#> Compute signaling network similarity for datasets 1 2
cellchat.merged <- netEmbedding(cellchat.merged, type = "functional")
#> Manifold learning of the signaling networks for datasets 1 2
cellchat.merged <- netClustering(cellchat.merged, type = "functional")
#> Classification learning of the signaling networks for datasets 1 2
# Visualization in 2D-space
p <- netVisual_embeddingPairwise(cellchat.merged, type = "functional", label.size = 3.5, do.label = T)
#> 2D visualization of signaling networks from datasets 1 2
#netVisual_embeddingZoomIn(cellchat.merged, type = "functional", nCol = 2)
ggsave("wt_functional.png",plot = p, device = "png", width = 20, height = 20)
```

Pathways show Ctrl_wt-OGD_wt
```{r}
pathways.show <- rev(unique(gg7[["data"]][(gg7[["data"]]$pvalues < 0.05) & ((gg7[["data"]]$contribution.relative < 
          1 - 0.05) | (gg7[["data"]]$contribution.relative > 1 + 0.05)),"name"]))

i <- 1
ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "outgoing", signaling = pathways.show, title = names(object.list)[i], width = 6, height = 12)
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "outgoing", signaling = pathways.show, title = names(object.list)[i+1], width = 6, height = 12)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))

ht3 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "incoming", signaling = pathways.show, title = names(object.list)[i], width = 6, height = 12, color.heatmap = "GnBu")
ht4 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "incoming", signaling = pathways.show, title = names(object.list)[i+1], width = 6, height = 12, color.heatmap = "GnBu")
draw(ht3 + ht4, ht_gap = unit(0.5, "cm"))

```

Pathways show Ctrl_w_a-OGD_w_a
```{r}
pathways.show <- rev(unique(gg7[["data"]][(gg7[["data"]]$pvalues < 0.05) & ((gg7[["data"]]$contribution.relative < 
          1 - 0.05) | (gg7[["data"]]$contribution.relative > 1 + 0.05)),"name"]))
```

Pathways show Ctrl_AxD-OGD_AxD
```{r}
pathways.show <- rev(unique(gg7[["data"]][(gg7[["data"]]$pvalues < 0.05) & ((gg7[["data"]]$contribution.relative < 
          1 - 0.05) | (gg7[["data"]]$contribution.relative > 1 + 0.05)),"name"]))
```
